### 2014-12-01 paper.tex
initial outline

\documentclass[12pt,letterpaper]{article}
\usepackage{url}
\usepackage{ctable}
\usepackage{overcite}
\usepackage[onehalfspacing]{setspace}
\usepackage{graphicx}
\usepackage{fullpage, times}

\DeclareGraphicsExtensions{.pdf,.png}
\onehalfspacing

\bibliographystyle{jcics}
\usepackage{anysize}

\begin{document}
\title{Exploring the Role of Small Molecules in Biological Systems Using Network Approaches}
\author{Saurav Das$^\dagger$ and Rajarshi Guha$^\ddagger$\\
$^\dagger$Your Address \\
$^\ddagger$National Center for Advancing Translational Science \\ 9800 Medical Center Drive  Rockville, MD 20850}
\date{}

\maketitle
\begin{abstract}
A nice abstract
\end{abstract}

\section{The Role of Networks in Drug Discovery}
\label{sec:role-networks-drug}

\section{Handling Small Molecules in R}
\label{sec:handl-small-molec}

\section{Linking Small Molecules to Targets, Pathways and Diseases}
\label{sec:link-small-molec}

\subsection{Drug-target networks}
\label{sec:drug-target-networks}

\subsection{Disease networks}
\label{sec:disease-networks}

\subsection{Dynamic networks}
\label{sec:dynamic-networks}

\section{R as a Platform for Computational Drug Discovery}
\label{sec:r-as-platform}

\section{Summary}
\label{sec:summary}

\end{document}

### 2014-12-02 paper.tex
Started section on R and cheminformatics

@@
 \usepackage{ctable}
-\usepackage{overcite}
 \usepackage[onehalfspacing]{setspace}
@@
 
+Small molecule representations are, in general, text based. However,
+due to the plethora of chemical formats, they are not natively handled
+within R.  Traditionally, one must compute numerical features for
+small molecules and perform cheminformatics related manipulation
+outside R and import the results of such operations into the R
+workspace. An alternative approach is to integrate cheminformatics
+toolkits such as the CDK \cite{Steinbeck:2003bh}, RDKit or Indigo into
+the R environment. Currently there are two packages that support this
+- ChemmineR \cite{Cao:2008fj} and rcdk \cite{Guha:2007aa}.
+
+In this section we briefly review the functionality provided by the
+rcdk package, primarily focused on manipulating molecular structures.
+
 \section{Linking Small Molecules to Targets, Pathways and Diseases}
@@
 
+\bibliography{paper}
+
 \end{document}

### 2014-12-29 paper.tex
Added intro for network representations of various entities

--- 2014-12-01 paper.tex
@@
 \usepackage{fullpage, times}
+\usepackage{listings}
+
+%% margin notes - comment out before submission
+\usepackage{marginnote}
+\usepackage[top=2.54cm, bottom=2.54cm, outer=5cm, inner=1.25cm, heightrounded, marginparwidth=4cm, marginparsep=0.25in]{geometry}
+\renewcommand*{\marginfont}{\color{red}\small}
 
@@
 \usepackage{anysize}
+
+\newcommand{\rcdk}{\texttt{rcdk}\ }
 
@@
 
+Though small molecule are represented in text based formats (SMILES,
+SDF, MOL2, InChI, etc.), the R environment does not support parsing
+and subsequent manipulation of small molecule structures. The
+traditional approach has been to compute numerical features for small
+molecules and perform cheminformatics related manipulation outside R
+and import the results of such operations into the R workspace. An
+alternative approach is to integrate cheminformatics toolkits such as
+the CDK \cite{Steinbeck:2003bh}, RDKit or Indigo into the R
+environment. Currently there are two packages that support this -
+\texttt{ChemmineR} \cite{Cao:2008fj} and \rcdk \cite{Guha:2007aa}. In this
+section we briefly review the functionality provided by the \rcdk
+package, which is an idiomatic R wrapper around the CDK Java library,
+primarily focused on manipulating molecular structures.
+
+Working with small molecule structure data can be broadly grouped into
+three tasks - input/output, manipulation \& modification of structures
+and computations on molecular structures. The \rcdk package supports
+input of all chemical file formats supported by the CDK, including
+SMILES, MOL MOL2, SDF and PDB formats. When used with the
+\texttt{rinchi}, input and output of the InChI format is also
+supported. Chemical structure files can be read locally or over the
+internet. In addition to generic file loading, helper functions are
+available for the SMILES format, given its ubiquitous use as a
+structure exchange format.
+
+Molecular structures are loaded in as references to Java objects. That
+is, they are not native R data structures and thus can only be
+manipulated using methods from the \texttt{rJava} package. While
+inconvenient, the low level details are hidden from view and the \rcdk
+package provides an idiomatic R interface to various CDK methods that
+operate on molecules, bonds and atoms. For example, given a SMILES
+string one can count the number of aromatic atoms using the following
+R code
+\begin{lstlisting}
+mol <- parse.smiles('CNCc1ccccc1')[[1]]
+length(which(sapply(get.atoms(mol), 
+    function(atom) 
+        is.aromatic(atom))))  
+\end{lstlisting}
+Other methods support operations on bonds, identifying substructures
+and retrieving atom or bond properties.
+
+\marginnote{A cmment}From an analytic point of view, functions that perform computations on
+molecular structures are probably the most useful. Such computations
+can range from evaluating molecular descriptors \cite{Guha:2012vn} and
+generating fingerprints to computing a variety of similarities and so
+on. The \rcdk package provides a simple interface to descriptor and
+fingerprint calculation. Importantly, the results from these functions
+can easily be employed in a network based approach when coupled with
+\texttt{igraph}. As an example, consider a similarity network
+constructed for 100 molecules, using a Tanimoto cutoff of 0.75
+
+\begin{lstlisting}
+library(rcdk)
+library(igraph)
+library(fingerprint)
+mols <- load.molecules('data/mipe-std.smi')
+fps <- lapply(mols, get.fingerprint)
+smat <- fp.sim.matrix(fps)
+smat[ smat < 0.75 ] <- 0
+g <- graph.adjacency(smat, mode='undirected', 
+    weight=TRUE, diag=FALSE)  
+\end{lstlisting}
+
 \section{Linking Small Molecules to Targets, Pathways and Diseases}
 \label{sec:link-small-molec}
+
+Since small molecules are fundamentally weighted networks (of atoms
+and bonds), a variety of graph algorithms can be applied to them to
+derive invariants (also known as topological descriptors
+\cite{Guha:2012vn}). However, small molecules interact with protein
+targets and the targets themselves interact with each other. Such
+explicit interactions lend themselves naturally to network
+representations. More generally, observed or computed relationships
+between small molecules, protein or gene targets and diseases allow
+one to develop network representations which can then be visualized
+and quantified to support integrative analyses of multiple data
+types. In the following subsections we highlight applications of this
+approach and where possible provide examples of R code to generate and
+analyze such networks.
 
@@
 \label{sec:disease-networks}
+
+\subsection{Assay networks}
+\label{sec:assay-networks}
+
+\subsection{Scaffold networks}
+\label{sec:scaffold-networks}
 

### 2014-12-29 code.R

@@ -1,0 +1,15 @@
+library(rcdk)
+library(igraph)
+library(fingerprint)
+mols <- load.molecules('data/mipe-std.smi')
+fps <- lapply(mols, get.fingerprint)
+smat <- fp.sim.matrix(fps)
+smat[ smat < 0.75 ] <- 0
+g <- graph.adjacency(smat, mode='undirected', weight=TRUE, diag=FALSE)
+
+par(mar=rep(0,4))
+pdf('img/sim-network.pdf', 6,6)
+plot(g, layout=layout.fruchterman.reingold,
+     margin=0,
+     vertex.label=NA, vertex.size=2)
+dev.off()

### 2014-12-30 paper.tex
first draft of assay network section

@@
 \usepackage{listings}
+\usepackage{hyperref}
 
@@
 
+Most high throughput screening campaigns tend to involve multiple
+assays. These usually include a primary screen followed by one or more
+secondary screens as well as counter-screens. When data from such
+campaigns are deposited into public databases such as Pubchem, the
+temporal sequence of the assays is not always maintained. As a result
+it can be difficult to track the progression of chemical matter
+through the series of assays. In addition, a view of the complete
+assay sequence can allow one to detect commonalities between programs
+and better compare the performance of screening workflows. Calhoun et
+al \cite{Calhoun:2012uq} describe an approach to reconstructing
+screening workflows (also termed workflow graphs) from screening
+datasets. They employed four heuristic rules to construct graphs from
+the data as well as using Pubchem Bioassay meta data and text mining
+to identify assays from a given project. These assay networks can be
+visualized using an online tool located at
+\href{http://swami.wustl.edu/flow}{http://swami.wustl.edu/flow}.
+Swamidass et al \cite{Swamidass:2014vn} describe an approach to
+constructing ``assay networks'' based on Pubchem data, where assays
+are connected when they exhibit common active (non-promiscuous)
+compounds. In this approach connected assays are correlated since they
+exhibit compounds with similar activities. This method leads to
+connections between apparently unrelated screen allowing one to
+identify compounds that exhibit polypharmacological behavior (i.e.,
+activity against multiple targets) and thus are candidates for
+repurposing. By also labeling screens as phenotypic or biochemical,
+this approach allows one to propose target candidates based on
+connectivity, thus serving as one approach to target deconvolution.
+
 \subsection{Scaffold networks}

### 2014-12-30 paper.tex
Started section on scaffold networks Added discussion points

@@
 
+The notion of a scaffold is common place in medicinal chemistry and is
+usually considered to be a ring-containing substructure pruned of
+sidechains that represents the core structure of a series of
+molecules. An example is the tricyclic core common to a number of
+antidepressants (e.g., amitriptylene and imipramine) shown in Figure
+\ref{fig:tca-core}. Scaffolds allow one to summarize a collection of
+molecules (such as in patent claims) and one can usually assume that
+compounds with the same scaffold will share a common synthetic pathway
+or even a common mechanism of action.
+
+While a number of approaches have been described to generate scaffolds
+\cite{Lewell:1998aa,Bemis:1996aa,Katritzky:2000wf}, a few approaches
+have considered hierarchical decomposition of scaffolds. Examples
+include the HierS method \cite{Wilkens:2005il} and the Scaffold Tree
+method \cite{Schuffenhauer:2007oz}.
+
+\ctable[cap={}, caption={An example of a scaffold. In this case, this
+  scaffold is common to many tricyclic antidepressants}, figure, botcap,
+label={fig:tca-core}]    
+{c}
+{}
+{\includegraphics{img/tca-core}}
+
 \subsection{Dynamic networks}
@@
 
-\section{Summary}
+\section{Discussion}
 \label{sec:summary}
 
+\begin{itemize}
+\item Networks on the chemistry side tend to be more for visualization
+\item In many cases the network does not necessarily lead to nove output
+\item A key utility of network approaches is the ability to integrate
+  different data types such as small molecules and proteins
+\item Networks represent an approach to representing multiple
+  relationships in a single contruct - multiple properties of a
+  relatinship between two entities can be encoded by edge properties
+  (width, color, length). Similarly multiple properties of a given
+  entity can be encoded in node properties (size, shape, color) 
+\end{itemize}
 \bibliography{paper}

### 2014-12-30 paper.tex
Fleshed out section on scaffold networks

@@
 or even a common mechanism of action.
-
-While a number of approaches have been described to generate scaffolds
-\cite{Lewell:1998aa,Bemis:1996aa,Katritzky:2000wf}, a few approaches
-have considered hierarchical decomposition of scaffolds. Examples
-include the HierS method \cite{Wilkens:2005il} and the Scaffold Tree
-method \cite{Schuffenhauer:2007oz}.
-
 \ctable[cap={}, caption={An example of a scaffold. In this case, this
@@
 {\includegraphics{img/tca-core}}
+
+While a number of approaches have been described to generate scaffolds
+\cite{Lewell:1998aa,Bemis:1996aa,Katritzky:2000wf}, a few approaches
+have considered hierarchical decomposition of scaffolds. Examples
+include the HierS method \cite{Wilkens:2005il} and the Scaffold Tree
+method \cite{Schuffenhauer:2007oz}. The latter generates a series of
+scaffolds based on an iterative removal of rings. The output of this
+method is a series of scaffolds arranged in a tree-like structure
+whose root node represents the simplest ring obtainable from the
+starting scaffold.  Two scaffolds are connected if there exists a
+parent--child relationship between the two. Scaffold trees generated
+by this method lend themselves naturally to network
+visualizations. Scaffold Hunter \cite{Wetzel:2009uq} is an example of
+a tool that employs a network representation of scaffold trees. Figure
+\ref{fig:scafftree} shows an example of a scaffold tree generated from
+a set of 602 pyruvate kinase inhibitors. The graph is colored based on
+$\log AC_{50}$ and the data is obtained from Pubchem AID 361. This
+visualization lends itself naturally to adding overlays such as
+molecular properties either on leaf nodes or else as aggregated values
+on non-leaf nodes. The scaffold tree has been applied in a number of
+studies \cite{Wetzel:2009uq,Renner:2009wm} and the structure of the
+tree can be used as a guide to exploring chemical space, focusing on
+regions (as represented by a scaffold) where there is minimal activity
+(a ``biological hole'') or high activity.
+
+\ctable[cap={}, caption={An example of a scaffold tree displayed as a
+  radial network, generated from a set of pyruvate kinase inhibitors.}, figure, botcap,
+label={fig:scafftree}]    
+{c}
+{}
+{\includegraphics[width=\linewidth]{img/scafftree-full}}
+
+Varin et al \cite{Varin:2011ve} describe ``scaffold networks'' which
+are similar in nature to scaffold trees but instead of a tree
+structure where each scaffold has a single parent, a DAG is
+constructed by considering all scaffolds at a given level of the
+hierarchy. An example of a scaffold network generated from Alosetron
+is shown in Figure \ref{fig:scaffnet}, where the blue structures are
+those that would be generated using the Scaffold Tree algorithm and
+the green structures would be the extra scaffolds generated using the
+Scaffold Network algorithm. Using this approach, together with the
+compound set enrichment method \cite{Varin:2010zh} the authors were
+able to efficiently identify active scaffold and active molecules
+
+\ctable[cap={}, caption={An example of the scaffolds generated using
+  the Scaffold Tree and Scaffold Network algorithms, starting from
+  Alosetron. Blue scaffolds are generated by both methods and green
+  scaffolds are generated only by the Scaffold Network
+  method. Modified from Figure 1 in Varin et al \cite{Varin:2011ve}.}, figure, botcap,
+label={fig:scaffnet}]    
+{c}
+{}
+{\includegraphics[width=0.75\linewidth]{img/scaffold-network}}
 
@@
 \item Networks represent an approach to representing multiple
-  relationships in a single contruct - multiple properties of a
-  relatinship between two entities can be encoded by edge properties
+  relationships in a single construct - multiple properties of a
+  relationship between two entities can be encoded by edge properties
   (width, color, length). Similarly multiple properties of a given

### 2015-01-03 paper.tex
Added another subsection
Added section on SAR networks

@@
 
+\subsection{SAR networks}
+\label{sec:sar-networks}
+
+Network representations have been used to characterize
+structure-activity relationship (SAR) datasets, most notably in the
+context of activity cliffs \cite{Maggiora:2006aa} - compounds that are
+structurally similar, but show very different activities. In the approach
+described by Guha and Van Drie \cite{Guha:2008aa}, a Structure
+Activity Landscape Index (SALI) value is computed for each pair of
+molecules in a dataset. The resultant matrix of SALI values is then
+visualized in a network form, where compounds (nodes) are connected by
+an edge if their SALI value lies above a user specified threshold
+(larger SALI values imply a bigger activity cliff). An example of such
+a network is shown in Figure \ref{fig:salinet}. The representation
+allows one to rapidly zoom in on compound pairs that exhibit activity
+cliffs. In addition, the network representation was also employed as a
+way to quantify the ability of a predictive model to correctly predict
+activity cliffs \cite{Guha:2008ab}.
+
+\ctable[cap={}, caption={A SALI network constructed from XXX. See
+  Ref.~\cite{Guha:2008aa} for further details}, figure, botcap,
+label={fig:salinet}]    
+{c}
+{}
+{\includegraphics[width=\linewidth]{img/salinet}}
+
+Network similarity graphs (NSG's) are an alternative network
+visualization of SAR data described by Wawer et al \cite{Wawer:2008aa}
+that employs a different characterization of activity cliffs. This
+approach is designed to multiplex information by combining
+connectivity (representing similarity relationships) with node color
+(potency) and size (discontinuity score \cite{Peltason:2007aa}). This
+approach has been implemented in the SARANEA tool
+\cite{Lounkine:2010fk} and extended \cite{Iyer:2011ij} the NSG concept
+to include mechanism of action information.
+
 \subsection{Assay networks}

### 2015-01-04 paper.tex
Added more text and refs to SAR networks

@@
 
+Networks have been applied to a number of other SAR related
+problems. For example, Webb et al \cite{Webb:2014tp} proposed a
+network approach to the interpretation of arbitrary machine learning
+models. Hanser et al \cite{Hanser:2014gl} constructed a network model
+to represent SAR knowledge data, resulting in a ``hypothesis
+network''. A key feature of this approach is the ability to integrate
+disparate data types and build predictive models on top of the
+network. Krein and Sukumar \cite{Krein:2011tt} describe an analysis of
+chemical spaces using a graph representation and compare chemical
+spaces using network metrics.
+
 \subsection{Assay networks}

### 2015-01-19 paper.tex
Getting started ....

@@
 \title{Exploring the Role of Small Molecules in Biological Systems Using Network Approaches}
-\author{Saurav Das$^\dagger$ and Rajarshi Guha$^\ddagger$\\
-$^\dagger$Your Address \\
+\author{Sourav Das$^\dagger$ and Rajarshi Guha$^\ddagger$\\
+$^\dagger$St. Jude Children's Research Hospital \\ 262 Danny Thomas Place, Memphis, TN 38105 \\
 $^\ddagger$National Center for Advancing Translational Science \\ 9800 Medical Center Drive  Rockville, MD 20850}

### 2015-01-19 paper.tex
Slight changes

@@
 \label{sec:role-networks-drug}
+Omics technologies are enabling us today to elucidate and interrogate complex relationships between proteins, genetic material and small molecules at an unprecedented speed and scale.
+Interrogation of relationships reveal common mechanism of action.
+This may result in a therapeutic candidate finding application in a seemingly unrelated disease that share a previously unknown mechanism of action.
+This type of relationship elucidation may also prevent unwanted side-effects if interrogation would reveal an unexpected link to a vital pathway that ought to remain unperturbed.
+The process of interrogation starts with experiments aimed at elucidating molecular mechanisms and pathways such as yeast two-hybrid techniques, protein and small molecule mass spectrometry, DNA microarrays and copy number variation studies.
+On the chemistry side, experiments may include high throughput screening of millions of chemical compounds against biological targets of therapeutic interest.
+These experiments result in pairwise interaction relationships from which the functional big picture biological network can then be elucidated \cite{Wu2010}.
+Community structure algorithm \cite{girvan2002community}, for example, allows one to discover a set of sparsely interconnected subsets of vertices, the subsets in turn however being densely connected.
+In the biological context, this allows connecting a set of biological cycles and smaller pathways into a big picture biological network from which inferences on new-disease application and cross-reactivity profile of a drug candidate can be drawn.
 

### 2015-01-19 paper.tex
More changes

@@
 Interrogation of relationships reveal common mechanism of action.
-This may result in a therapeutic candidate finding application in a seemingly unrelated disease that share a previously unknown mechanism of action.
+This may lead to a therapeutic candidate finding application in a seemingly unrelated disease that share a previously unknown causal mechanism.
+It also allows identification of drug pairs that show synergistic activity stemming from their interaction with complementary pathways.
 This type of relationship elucidation may also prevent unwanted side-effects if interrogation would reveal an unexpected link to a vital pathway that ought to remain unperturbed.
 The process of interrogation starts with experiments aimed at elucidating molecular mechanisms and pathways such as yeast two-hybrid techniques, protein and small molecule mass spectrometry, DNA microarrays and copy number variation studies.
-On the chemistry side, experiments may include high throughput screening of millions of chemical compounds against biological targets of therapeutic interest.
+Experiments may also include high throughput screening of millions of chemical compounds against specific biological targets of therapeutic interest to find respective agonists and antagonists.
 These experiments result in pairwise interaction relationships from which the functional big picture biological network can then be elucidated \cite{Wu2010}.
-Community structure algorithm \cite{girvan2002community}, for example, allows one to discover a set of sparsely interconnected subsets of vertices, the subsets in turn however being densely connected.
-In the biological context, this allows connecting a set of biological cycles and smaller pathways into a big picture biological network from which inferences on new-disease application and cross-reactivity profile of a drug candidate can be drawn.
+Community structure algorithm \cite{girvan2002community} may be used, that allows one to discover a set of sparsely interconnected subsets of vertices, the subsets in turn however being densely connected within themselves.
+In the biological context, this allows connecting a set of biological cycles and smaller pathways into a big picture biological metacommunity network from which inferences on cross-reactivity and new application areas of a drug candidate can be drawn.
+
+Network based approach to mechanism elucidation of small molecules that allow identification of synergistic pairs have been implemented in tools such as the DrugComboRanker \cite{huang2014drugcomboranker}.
+The authors use a community based partitioning to create a drug metacommunity network and a disease signaling metacommunity network.
+Drug targets are then mapped onto disease networks to allow identification of synergistic pairs of drugs and complementary mechanism of action.
 

### 2015-01-19 paper.tex
Updated Role of Networks in Drug Discovery

@@
 \label{sec:role-networks-drug}
-Omics technologies are enabling us today to elucidate and interrogate complex relationships between proteins, genetic material and small molecules at an unprecedented speed and scale.
+Omics technologies are enabling us today to elucidate and interrogate complex relationships between disease, proteins, genetic material and small molecules at an unprecedented speed and scale.
 Interrogation of relationships reveal common mechanism of action.
-This may lead to a therapeutic candidate finding application in a seemingly unrelated disease that share a previously unknown causal mechanism.
-It also allows identification of drug pairs that show synergistic activity stemming from their interaction with complementary pathways.
+This may lead to a therapeutic candidate finding application in a seemingly unrelated disease (repositioning) that share a previously unknown causal mechanism.
+It also allows identification of drug pairs that show synergistic activity stemming from their interactions with complementary pathways.
 This type of relationship elucidation may also prevent unwanted side-effects if interrogation would reveal an unexpected link to a vital pathway that ought to remain unperturbed.
-The process of interrogation starts with experiments aimed at elucidating molecular mechanisms and pathways such as yeast two-hybrid techniques, protein and small molecule mass spectrometry, DNA microarrays and copy number variation studies.
+The process of interrogation starts with experiments aimed at elucidating molecular mechanisms and pathways such as yeast two-hybrid techniques, protein and small molecule mass spectrometry, microarray and copy number variation studies.
 Experiments may also include high throughput screening of millions of chemical compounds against specific biological targets of therapeutic interest to find respective agonists and antagonists.
 These experiments result in pairwise interaction relationships from which the functional big picture biological network can then be elucidated \cite{Wu2010}.
-Community structure algorithm \cite{girvan2002community} may be used, that allows one to discover a set of sparsely interconnected subsets of vertices, the subsets in turn however being densely connected within themselves.
+Algorithms that detect communities from complex networks eg.
+community structure algorithm \cite{girvan2002community} may be used.
+These network analyses algorithms allow discovery of a set of sparsely interconnected subsets of vertices where the subsets themselves are however densely connected within.
 In the biological context, this allows connecting a set of biological cycles and smaller pathways into a big picture biological metacommunity network from which inferences on cross-reactivity and new application areas of a drug candidate can be drawn.
@@
 The authors use a community based partitioning to create a drug metacommunity network and a disease signaling metacommunity network.
-Drug targets are then mapped onto disease networks to allow identification of synergistic pairs of drugs and complementary mechanism of action.
+Drug targets are then mapped onto disease networks to allow identification of synergistic pairs of drugs and complementary mechanisms of action.
+In the tool MANTRA (Mode of Action by NeTwoRk Analysis)\cite{iorio2010discovery}, authors are able to deduce mechanisms of action of new molecules and suggest repositioning of a well known molecule to a new area of application.
+
+Network analysis has also found its application in Medicinal Chemistry and High-throughput Screening.
+Cheminformatics based chemical scaffold and topology networks capture the structure-activity relationships between compounds in screening experiments \cite{varin2011mining}.
+These help in medicinal chemistry efforts towards the identification and optimization of lead compounds for the  biological target of interest by allowing visualization of the chemical space in the backdrop of biological activity.
 

### 2015-01-19 paper.tex
Added wiley style file and refactored tex doc to use Wiley style. Means we have to drop use of ctable

@@
-\documentclass[12pt,letterpaper]{article}
-\usepackage{url}
-\usepackage{ctable}
-\usepackage[onehalfspacing]{setspace}
+\documentclass[]{book}
+
+\usepackage{listings}
+
+\usepackage[utf8]{inputenc}
 \usepackage{graphicx}
-\usepackage{fullpage, times}
-\usepackage{listings}
-\usepackage{hyperref}
+\usepackage[]{natbib}
+\bibliographystyle%
+  {wileynum}%for numerical citation and numerically listed entries in the bibliography
+  %{wileyauy}%for author--year citation and alphabetical order in the bibliography
+\usepackage{wileySTM}
+\usepackage[\ifnum\pdfoutput=1breaklinks\fi]{hyperref}
 
-%% margin notes - comment out before submission
-\usepackage{marginnote}
-\usepackage[top=2.54cm, bottom=2.54cm, outer=5cm, inner=1.25cm, heightrounded, marginparwidth=4cm, marginparsep=0.25in]{geometry}
-\renewcommand*{\marginfont}{\color{red}\small}
+\title{Exploring the Role of Small Molecules in Biological Systems Using Network Approaches}
+\author{Sourav Das and Rajarshi Guha}
+
+% \author{Sourav Das$^\dagger$ and Rajarshi Guha$^\ddagger$}\\
+% $^\dagger$St. Jude Children's Research Hospital \\ 262 Danny Thomas Place, Memphis, TN 38105 \\
+% $^\ddagger$National Center for Advancing Translational Science \\ 9800 Medical Center Drive  Rockville, MD 20850}
 
 \DeclareGraphicsExtensions{.pdf,.png}
-\onehalfspacing
-
-\bibliographystyle{jcics}
-\usepackage{anysize}
 
@@
 \begin{document}
-\title{Exploring the Role of Small Molecules in Biological Systems Using Network Approaches}
-\author{Sourav Das$^\dagger$ and Rajarshi Guha$^\ddagger$\\
-$^\dagger$St. Jude Children's Research Hospital \\ 262 Danny Thomas Place, Memphis, TN 38105 \\
-$^\ddagger$National Center for Advancing Translational Science \\ 9800 Medical Center Drive  Rockville, MD 20850}
-\date{}
-
-\maketitle
-\begin{abstract}
-A nice abstract
-\end{abstract}
+\frontmatter
+\mainmatter
 
@@
 
-\marginnote{A cmment}From an analytic point of view, functions that perform computations on
+From an analytic point of view, functions that perform computations on
 molecular structures are probably the most useful. Such computations
@@
 
-\ctable[cap={}, caption={A SALI network constructed from XXX. See
-  Ref.~\cite{Guha:2008aa} for further details}, figure, botcap,
-label={fig:salinet}]    
-{c}
-{}
-{\includegraphics[width=\linewidth]{img/salinet}}
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=\linewidth]{img/salinet} 
+  \caption{A SALI network constructed from XXX. See Ref.~\cite{Guha:2008aa} for further details.}
+  \label{fig:salinet}
+\end{figure}
 
@@
 or even a common mechanism of action.
-\ctable[cap={}, caption={An example of a scaffold. In this case, this
-  scaffold is common to many tricyclic antidepressants}, figure, botcap,
-label={fig:tca-core}]    
-{c}
-{}
-{\includegraphics{img/tca-core}}
+\begin{figure}[h]
+  \centering
+  \includegraphics{img/tca-core}
+  \caption{An example of a scaffold. In this case, this
+  scaffold is common to many tricyclic antidepressants.}
+  \label{fig:tca-core}
+\end{figure}
 
@@
 
-\ctable[cap={}, caption={An example of a scaffold tree displayed as a
-  radial network, generated from a set of pyruvate kinase inhibitors.}, figure, botcap,
-label={fig:scafftree}]    
-{c}
-{}
-{\includegraphics[width=\linewidth]{img/scafftree-full}}
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.9\linewidth]{img/scafftree-full}
+  \caption{An example of a scaffold tree displayed as a  radial
+    network, generated from a set of pyruvate kinase inhibitors.}
+  \label{fig:scafftree}
+\end{figure}
 
@@
 
-\ctable[cap={}, caption={An example of the scaffolds generated using
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.75\linewidth]{img/scaffold-network}
+  \caption{An example of the scaffolds generated using
   the Scaffold Tree and Scaffold Network algorithms, starting from
@@
   scaffolds are generated only by the Scaffold Network
-  method. Modified from Figure 1 in Varin et al \cite{Varin:2011ve}.}, figure, botcap,
-label={fig:scaffnet}]    
-{c}
-{}
-{\includegraphics[width=0.75\linewidth]{img/scaffold-network}}
+  method. Modified from Figure 1 in Varin et al \cite{Varin:2011ve}.}
+  \label{fig:scaffnet}
+\end{figure}
 

### 2015-02-02 paper.tex
some more text

@@
 
+The core concepts involved in computational drug discovery can be
+reduced to genes, proteins, pathways and small molecules interacting
+with them. Computational representations of these concepts and the
+methods to be manipulate them have been implemented in a variety of
+software tools. A comprehensive computational drug discovery platform
+will include a number of these tools, as it is rare that a single tool
+addresses all aspects.
+
+In this sense, the R environment presents a robust platform for
+computational drug discovery, especially the stages up to and
+including lead optimization. The Bioconductor set of packages provide
+a plethora of tools to analyze a variety of ``omics'' data. Examples
+include packages that support gene expression (microarray or RNAseq)
+analyses, tools for differential expression analysis, gene and pathway
+enrichment analysis and so on.
+
+In contrast, support for small molecule data is limited. Currently,
+there are two packages that focus on representing and manipulating
+chemical structures - \texttt{ChemmineR} \cite{Cao:2008fj} and \rcdk
+\cite{Guha:2007aa}. Both expose cheminformatics functionality
+implemented in an underlying Java library (JOELib for ChemmineR and
+the CDK for \rcdk). In this section we focus on the use of \rcdk as a
+means to manipulate and analyze chemical structure data.
+
+The \rcdk package supports a variety of chemical structure file
+formats and presents the CDK Java API as a set of idiomatic R
+functions. That is, rather than manually specify fully qualified Java
+packages or resort to writing ``Java in R'' the \rcdk package wraps
+much of the CDK API as ordinary R functions. Thus the atoms in a
+molecule are represented as a set of atom objects. Importantly, atoms
+(and other objects such as bonds and molecules) are stored as
+references to Java objects. While they can be manipulated directly, it
+is unwieldy it is recommended that one uses methods provided by the
+\rcdk package.
+
 \section{Discussion}

### 2015-02-23 paper.tex
Moved and updated section on using R for network analysis in drug discovery

@@
 \newcommand{\rcdk}{\texttt{rcdk}\ }
+\newcommand{\igraph}{\texttt{igraph}\ }
 
@@
 \mainmatter
+
+\chapter{Exploring the Role of Small Molecules in Biological Systems Using Network Approaches}
 
@@
 These help in medicinal chemistry efforts towards the identification and optimization of lead compounds for the  biological target of interest by allowing visualization of the chemical space in the backdrop of biological activity.
-
-\section{Handling Small Molecules in R}
-\label{sec:handl-small-molec}
-
-Though small molecule are represented in text based formats (SMILES,
-SDF, MOL2, InChI, etc.), the R environment does not support parsing
-and subsequent manipulation of small molecule structures. The
-traditional approach has been to compute numerical features for small
-molecules and perform cheminformatics related manipulation outside R
-and import the results of such operations into the R workspace. An
-alternative approach is to integrate cheminformatics toolkits such as
-the CDK \cite{Steinbeck:2003bh}, RDKit or Indigo into the R
-environment. Currently there are two packages that support this -
-\texttt{ChemmineR} \cite{Cao:2008fj} and \rcdk \cite{Guha:2007aa}. In this
-section we briefly review the functionality provided by the \rcdk
-package, which is an idiomatic R wrapper around the CDK Java library,
-primarily focused on manipulating molecular structures.
-
-Working with small molecule structure data can be broadly grouped into
-three tasks - input/output, manipulation \& modification of structures
-and computations on molecular structures. The \rcdk package supports
-input of all chemical file formats supported by the CDK, including
-SMILES, MOL MOL2, SDF and PDB formats. When used with the
-\texttt{rinchi}, input and output of the InChI format is also
-supported. Chemical structure files can be read locally or over the
-internet. In addition to generic file loading, helper functions are
-available for the SMILES format, given its ubiquitous use as a
-structure exchange format.
-
-Molecular structures are loaded in as references to Java objects. That
-is, they are not native R data structures and thus can only be
-manipulated using methods from the \texttt{rJava} package. While
-inconvenient, the low level details are hidden from view and the \rcdk
-package provides an idiomatic R interface to various CDK methods that
-operate on molecules, bonds and atoms. For example, given a SMILES
-string one can count the number of aromatic atoms using the following
-R code
-\begin{lstlisting}
-mol <- parse.smiles('CNCc1ccccc1')[[1]]
-length(which(sapply(get.atoms(mol), 
-    function(atom) 
-        is.aromatic(atom))))  
-\end{lstlisting}
-Other methods support operations on bonds, identifying substructures
-and retrieving atom or bond properties.
-
-From an analytic point of view, functions that perform computations on
-molecular structures are probably the most useful. Such computations
-can range from evaluating molecular descriptors \cite{Guha:2012vn} and
-generating fingerprints to computing a variety of similarities and so
-on. The \rcdk package provides a simple interface to descriptor and
-fingerprint calculation. Importantly, the results from these functions
-can easily be employed in a network based approach when coupled with
-\texttt{igraph}. As an example, consider a similarity network
-constructed for 100 molecules, using a Tanimoto cutoff of 0.75
-
-\begin{lstlisting}
-library(rcdk)
-library(igraph)
-library(fingerprint)
-mols <- load.molecules('data/mipe-std.smi')
-fps <- lapply(mols, get.fingerprint)
-smat <- fp.sim.matrix(fps)
-smat[ smat < 0.75 ] <- 0
-g <- graph.adjacency(smat, mode='undirected', 
-    weight=TRUE, diag=FALSE)  
-\end{lstlisting}
 
@@
 
-\section{R as a Platform for Computational Drug Discovery}
+\section{R as a Platform for Network Analyses in Drug Discovery}
 \label{sec:r-as-platform}
@@
 
+For a detailed review of \rcdk functionality, the reader is referred
+to Guha \cite{Guha:2007aa}. For now we focus on simple network
+analyses that involve chemical structures. Given a set of compounds
+represented as SMILES strings, \rcdk can be used to parse them,
+compute binary fingerprints and then compute a pairwise Tanimoto
+similarity matrix, as shown below
+\begin{lstlisting}
+library(rcdk)
+library(fingerprint)
+
+mols <- parse.smiles(mipe$SMILES_STD)
+fps <- lapply(mols, get.fingerprint, type='extended')
+fpsim <- fp.sim.matrix(fps)
+colnames(fpsim) <- mipe$SAMPLE_NAME
+rownames(fpsim) <- mipe$SAMPLE_NAME
+fpsim[ fpsim < 0.6 ] <- 0
+\end{lstlisting}
+In this code snippet, we compute a pairwise similarity matrix that we
+will eventually use as an adjacency matrix to construct the similarity
+network. We only keep those similarity values greater than 0.6 so as
+to focus only on similar molecules.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.75\linewidth]{img/sim-network-connected}
+  \caption{An example of a similarity network, where nodes are
+    molecules and are connected if they exhibit a Tanimoto similarity
+    greater than 0.6. Nodes are colored based on the PANTHER
+    \cite{Mi:2005qq} class of their primary target.}
+  \label{fig:simnet}
+\end{figure}
+
+The resultant similarity matrix can be treated as an adjacency matrix
+and used as input to \igraph and subsequently plotted
+\begin{lstlisting}
+g <- graph.adjacency(fpsim, mode='undirected', 
+                     weight=TRUE, diag=FALSE)
+plot(g, layout=layout.fruchterman.reingold)
+\end{lstlisting}
+As example of such an analysis is shown in Figure \ref{fig:simnet}. In
+this network, we consider a set of 1,341 compounds annotated with a
+primary gene target. For each target, the highest level PANTHER
+\cite{Mi:2005qq} classification was determined and used to color the
+nodes. In addition, we ignore compounds that were less than 0.6
+similar to any other compound (i.e., singleton nodes).
+
+This is purely a visualization of a similarity matrix,
+modified to focus on similar (according to a specified threshold)
+compounds. Importantly, we have been to overlay a second type of
+information (i.e., target class) on top of the similarity values and
+is thus an example of the integrative nature of a network
+representation.
+
+Given the compound network and the associated target information, one
+could easily identify protein-protein interactions (BioGRID
+\cite{Chatr-Aryamontri:2015yf}, STRING \cite{Franceschini:2013qa} and
+other sources) and generate a bipartite graph of compounds and protein
+targets. Even in the trivial, annotated network show in Figure
+\ref{fig:simnet} it is evident that chemical similarity is related to
+functional similarity of the compounds' primary target - compounds
+targeting cytoskeletal proteins tend to be similar to each other;
+though interestingly, within the set of compounds targeting
+cytoskeletal proteins there appear to be five distinct groupings. When
+coupled to an interactive platform, this visualization can support
+rapid identification and exploration of novel associations.
+
 \section{Discussion}
@@
 
-\begin{itemize}
-\item Networks on the chemistry side tend to be more for visualization
-\item In many cases the network does not necessarily lead to nove output
-\item A key utility of network approaches is the ability to integrate
-  different data types such as small molecules and proteins
-\item Networks represent an approach to representing multiple
-  relationships in a single construct - multiple properties of a
-  relationship between two entities can be encoded by edge properties
-  (width, color, length). Similarly multiple properties of a given
-  entity can be encoded in node properties (size, shape, color) 
-\end{itemize}
+In this chapter we have attempted to provide an overview of how
+networks have been employed to analyze and visualize chemical
+data. The fact that chemical structures themselves are graphs allows
+one to apply a number of techniques from graph theory to characterize
+chemical structure (such as the many graph invariants). However, with
+the advent of large collections of chemical structures (from databases
+such as Pubchem or high throughput screening programs) along with
+associated data such as bioactivity and target annotations, a number
+of workers have attempted to analyze the properties of such
+collections using network techniques.
+
+It is interesting to note that many applications of network methods to
+chemical collections focus on visualization of the chemical data,
+rather than using the network structure to draw insight into the
+underlying chemical of biological phenomena. This is not to say that
+network visualizations of chemical data are lacking. In fact, such
+visualizations allow one to encode multiple properties of objects
+(i.e., nodes) and relationships (i.e., edges) in a single visual form.
+
+However, a key utility of network approaches is the ability to
+integrate different data types (such as small molecules and their
+protein targets) in to a single data structure. Drug-target or
+drug-disease (or even drug-disease-target) networks are examples of
+such approaches. Hartsperger et al \cite{Hartsperger:2010yg} describe
+an approach to fuzzy clustering that is capable of identifying
+subgraphs with heterogenous networks. A characteristic feature of such
+$k$-partite networks is that it allows one to model the effects of
+each class of entity on other classes of entities. This is exemplified
+by random walk approaches on bipartite networks such as that described
+by Chen et al \cite{Chen:2012qy} whose proposed a random walk approach
+to predict drug-target interactions.
+
 \bibliography{paper}

### 2015-02-23 code.R

@@ -1,15 +1,83 @@
 library(rcdk)
 library(igraph)
 library(fingerprint)
-mols <- load.molecules('data/mipe-std.smi')
-fps <- lapply(mols, get.fingerprint)
-smat <- fp.sim.matrix(fps)
-smat[ smat < 0.75 ] <- 0
-g <- graph.adjacency(smat, mode='undirected', weight=TRUE, diag=FALSE)
 
-par(mar=rep(0,4))
-pdf('img/sim-network.pdf', 6,6)
+library(org.Hs.eg.db)
+library(DBI)
+library(RMySQL)
+
+m<-dbDriver("MySQL");
+con<-dbConnect(m,user='guhar',password='beeb1e',host='bohr.ncats.nih.gov',dbname='bard3');
+
+mipe <- read.csv('data/mipe.csv', header=TRUE, as.is=TRUE)
+mipe <- subset(mipe, GENE_SYMBOL != '')
+
+## Convert syms to uniprot ID's and then pull in Panther class
+## but only keep the first Panther class at node_level = 1
+eids <- select(org.Hs.eg.db, keys=mipe$GENE, columns=c("SYMBOL", "ENTREZID", "UNIPROT"), keytype="SYMBOL")
+sql <- "select a.pclass_id, node_level, class_name from panther_uniprot_map a, panther_class b where accession in (%s) and a.pclass_id = b.pclass_id"
+tmp <- by(eids, eids$SYMBOL, function(x) {
+  accs <- join(sprintf("'%s'", x$UNIPROT), ',')
+  q <- sprintf(sql, accs)
+  r <- dbGetQuery(con, q)
+  if (nrow(r) == 0) ret <- data.frame(gene=x$SYMBOL[1], pclass_id=NA, node_level=NA, class_name="UNKNOWN")
+  else {
+    ret <- cbind(gene=x$SYMBOL[1], r)
+    ret <- subset(ret, node_level == 1)[1,]
+  }
+  return(ret)
+})
+tmp <- do.call(rbind, tmp)
+tmp.mipe <- merge(mipe, tmp, by.x='GENE_SYMBOL', by.y='gene')
+mipe <- tmp.mipe
+
+mols <- parse.smiles(mipe$SMILES_STD)
+fps <- lapply(mols, get.fingerprint, type='extended')
+fpsim <- fp.sim.matrix(fps)
+colnames(fpsim) <- mipe$SAMPLE_NAME
+rownames(fpsim) <- mipe$SAMPLE_NAME
[…]

### 2015-02-23 code.R
dropped a section

@@ -69,7 +69,7 @@ dev.off()
 non.singles <- which(degree(g) > 0)
 g2 <- induced.subgraph(g, vids = non.singles)
 vcols <- unlist(sapply(V(g2)$panther, function(x) col21[x]))
-pdf('img/sim-network-connected.pdf', 8,7)
+pdf('img/sim-network-connected.pdf', 9,7)
 layout(matrix(c(1,2,1,2), 2,2, byrow=TRUE), widths=c(3,1), heights=c(1,1))
 par(mar=c(0,0,0,0))
 plot(g2, layout=layout.fruchterman.reingold,
@@ -78,6 +78,6 @@ plot(g2, layout=layout.fruchterman.reingold,
 par(mar=c(0,0,0,0.5))
 plot.new()
 legend('center', legend=names(col21),  fill=unlist(col21),
-       border='white', pt.cex=8, bty='n', y.intersp=1.5,
-       title='PANTHER Class')
+       border='white', cex=1.2, pt.cex=8, bty='n', y.intersp=1.5,
+       title=expression(bold('PANTHER Class')))
 dev.off()

### 2015-02-23 paper.tex
dropped a section
Added section on disease networks. Added some config for listings

@@
 \usepackage{listings}
+\lstset{% general command to set parameter(s)
+basicstyle=\small, % print whole listing small
+keywordstyle=\color{red}\itshape,
+% underlined bold black keywords
+commentstyle=\color{blue}, % white comments
+stringstyle=\ttfamily, % typewriter type for strings
+showstringspaces=false,
+numbers=left, numberstyle=\tiny, stepnumber=1, numbersep=5pt, %
+rulesepcolor=\color{black}
+}
 
@@
 \label{sec:disease-networks}
+
+Schadt et al \cite{Schadt:2009zl} have stressed the importance of
+viewing diseases as manifestations of perturbed biological networks
+and consequently, the importance of drug discovery approaches that
+address the networks underlying a disease state.. Network approaches
+to exploring diseases have traditionally explored biological networks
+(e.g., protein-protein interaction or gene regulatory networks) in the
+context of individual diseases \cite{Lim:2006fe,
+  Altieri:2008aa}. Other approaches on the other hand have focused on
+classes of disease \cite{Zhang:2013tx} or even all known human
+diseases \cite{Goh:2007aa} (the so called ``diseaseome''). While most
+of these studies have primarily focused on identifying relationships
+between diseases in different ways such as sharing implicated genes
+\cite{Bauer-Mehren:2011tv,Goh:2007aa}, sharing symptoms
+\cite{Zhou:2014qp} or based on bibliometric associations
+\cite{Zhang:2014sh}. Studies on disease networks tend to construct
+heterogenous (or $k$-partite) networks where nodes can represent
+different classes of entities. An example is the network constructed
+from long non-coding RNAs and diseases by Zhou et al
+\cite{Zhou:2015if}. Based on this network, a random walk approach was
+employed to identify novel lncRNA-disease associations. These
+approaches play an important role in drug discovery as they provide
+opportunities for repurposing drugs based on novel disease-target or
+disease-pathway associations. Another view provided by such analysis
+is the druggability of a disease; that is, based on underlying targets
+and pathways, is it even feasible to treat the disease using a small
+molecule intervention \cite{Boran:2010fk}. An important function of
+network analyses of diseases is to go beyond simply identifying new
+targets for diseases or links between diseases, but instead attempt to
+elucidate the mechanisms underlying condition. As an example of this is
+the paradoxical effects of MEK inhibition and PI3K/AKT inhibition and
+the effects on cancer resistance \cite{Sos:2009gc}. Together with knowledge
+of the underlying network connectivity, these approaches allow one to
+design combination treatments - multiple drugs that affect multiple
+targets and pathways. This approach was exemplified by Lehar et al
+\cite{Lehar:2009gu} who proposed that drug combination response
+surfaces were functionally related to the underlying connectivity of
+the drug targets.
 
@@
 
-\subsection{Dynamic networks}
-\label{sec:dynamic-networks}
-
 \section{R as a Platform for Network Analyses in Drug Discovery}

### 2015-03-01 paper.tex
ensure proper figure size
updated author ordering Started on intro section

@@
 \title{Exploring the Role of Small Molecules in Biological Systems Using Network Approaches}
-\author{Sourav Das and Rajarshi Guha}
+\author{Rajarshi Guha and Saurav Das}
 
@@
 \chapter{Exploring the Role of Small Molecules in Biological Systems Using Network Approaches}
+
+The traditional, reductionist view of biology has considered the
+behavior and interactions of individual molecular entities as a way to
+understand and explain biological phenomena. This approach has become
+embodied in the ``one-target, one-drug''  approach to drug
+development. Leaving aside the subtleties of how targets are defined,
+it has become apparent that this model does not correspond very well
+to reality. Instead, it has become increasingly clear that many drugs
+(and more generally, small molecules) interact with multiple targets,
+including their putative primary target.
 
@@
   \centering
-  \includegraphics[width=0.75\linewidth]{img/sim-network-connected}
+  \includegraphics[width=\linewidth]{img/sim-network-connected}
   \caption{An example of a similarity network, where nodes are

### 2015-03-01 scafdoc.R
Added code to generate scaffold-doc networks

@@ -1,0 +1,16 @@
+library(igraph)
+
+id <- 6687479
+g <- read.graph(sprintf('4-%d.gml',id), format='gml')
+
+## filter out singletons
+#singleton.idx <- which(degree(g) > 0)
+#g <- induced.subgraph(g, vids=singleton.idx)
+
+vcol <- rep('black', length(V(g)))
+vcol[ which(V(g)$klass == 'doc') ] <- 'red'
+
+pdf(sprintf("%d.pdf", id), 20,20)
+par(mar=rep(0,4))
+plot(g, layout=layout.fruchterman.reingold, vertex.size=2, vertex.label=NA, vertex.color=vcol)
+dev.off()

### 2015-03-01 scafdoc.py

@@ -1,0 +1,122 @@
+import MySQLdb
+from igraph import *
+from igraph import _igraph
+import Queue
+
+global n
+max_round = 4
+
+db = MySQLdb.connect(host="carnot.ncats.nih.gov", # your host, usually localhost
+                     user="chembl_18", # your username
+                      passwd="chembl_18", # your password
+                      db="chembl_18") # name o
+cur = db.cursor()
+
+def node_exists(g,name):
+    try:
+        g.vs.find(Name=name)
+        return True
+    except:
+        return False
+    
+def pg(g):
+    color_dict = {"doc": "red", "fragment": "black"}
+    g.vs["color"] = [color_dict[klass] for klass in g.vs["klass"]]
+    g.vs['label'] = g.vs['Name']
+    layout = g.layout("fr")
+    
+    plot(g, layout = layout, vertex_size=5)
+
+def get_docs_for_fragid(fid, min_year = 1997, max_year = 2015, cur = None):
+    sql = """select distinct pubmed_id, title, abstract, year
+  from ncgc_fragment_class x, ncgc_fragment_instances a, activities b, docs c
+where x.class_id = %d
+and x.class_id = a.class_id
+and a.molregno = b.molregno
+and c.doc_id = b.doc_id
+and year >=  %d and year <= %d
+and pubmed_id is not null""" % (fid, min_year, max_year)
+
+    cur.execute(sql)
+    return cur.fetchall()
+
+def get_frag_for_docid(pmid, max_inst = 300, min_inst = 20, min_complex = 300,  cur = None):
+    sql = """select distinct x.class_id, x.smiles, acount, complexity, symmetry, instances
+  from ncgc_fragment_class x, ncgc_fragment_instances a, activities b, docs c
+where c.pubmed_id = %d
+and c.doc_id = b.doc_id
+and a.molregno = b.molregno
+and x.class_id = a.class_id and instances < %d and instances > %d and complexity > %d""" % (pmid, max_inst, min_inst, min_complex)
+    cur.execute(sql)
+    return cur.fetchall()
+
+def handle_doc_nodes(g, node, new_nodes):
+    assert node['klass'] == 'fragment'
+    node_name = node['Name']
+    for pmid, title, abstract, year in new_nodes:
+        try:
+            eid = g.get_eid( g.vs.find(Name=node_name).index, g.vs.find(Name=pmid), directed=False, error=True )
[…]

### 2015-03-02 paper.tex
Added section on scaffold doc network

@@
 
-\chapter{Exploring the Role of Small Molecules in Biological Systems Using Network Approaches}
+\chapter{Exploring the Role of Small Molecules in Biological Systems
+  Using Network Approaches}
 
@@
 development. Leaving aside the subtleties of how targets are defined,
-it has become apparent that this model does not correspond very well
-to reality. Instead, it has become increasingly clear that many drugs
-(and more generally, small molecules) interact with multiple targets,
-including their putative primary target.
+it has become apparent that this model does not always match
+reality. Instead, it has become increasingly clear that many drugs
+(and more generally, small molecules) exhibit polypharmacological
+behavior - interacting with multiple targets, including their putative
+primary target. In many cases, the activity against the primary target
+is significantly higher than off-target activities, minimizing the
+effects of the latter. However, this is not always the case as in the
+example of terfenadine (a histamine H1 antagonist), where off target
+activity against hERG led to an increased risk of cardiac death. But
+more importantly, it is becoming increasingly clear that many diseases
+are not driven by dysfunction of a single a target. Examples include
+diabetes and various cancers. In parallel, even when a single target
+is the main driver for a given disease state, compensatory pathways
+can ``route around'' inhibition of the primary target, thus restoring
+the functional behavior that the primary target is responsible (i.e.,
+the system is now resistant to the original drug)
+\cite{Sierra:2010mz,Dienstmann:2012qr}. Finally, it should be noted
+that while the preceding discussion presents polypharmacological
+behavior as a negative characteristic, this is not always the
+case. Indeed, modern approaches to drug development now consider such
+behavior attractive leading to efforts to rationally design or control
+polypharmacology \cite{Gujral:2014dp,Ciceri:2014fv,Metz:2010ys}.
+
+Given these observations it is clear that pathways play fundamental
+roles in biological systems. Naturally, approaches that analyse these
+networked systems can be effective in understanding their behavior and
+then identifying ways to modulate them. The remainder of this chapter
+looks at how network methods can be applied to various topics in drug
+discovery ranging from drug-target interactions to analyzing
+structure-activity relationships.
 
@@
 
+\subsection{Scaffold-document networks}
+\label{sec:scaff-docum-netw}
+
+With the advent of SAR databases such as ChEMBL \cite{Gaulton:2012jl},
+it is feasible to explore chemical space from a bibliometric point of
+view. For example, ChEMBL has identified small molecules mentioned in
+publications and provides the links between them. As a result, given a
+set of document it is possible to characterize the chemical space
+represented in those documents. As noted in Section
+\ref{sec:scaffold-networks}, scaffolds (or frameworks) allow one to
+better capture the core structures representing multiple
+compounds. Thus, a publication might refer to ten compounds from a
+single series, in which case it refers to a single scaffold (though
+this could depend on how one defines a scaffold). To explore this
+aspect of a SAR database, we applied a fragment decomposition method
+to all compounds reported in ChEMBL and linked the fragments
+associated with each molecule to their corresponding documents. The
+resultant data can then be represented as a bipartite network of
+fragments and documents which are connected when a document contains
+molecules that have the fragment as a substructure.
+
+One application of this dataset is to explore the nature of fragments
+associated with a document, and for each of those fragments identify
+the documents associated with them. This procedure can be repeated to
+construct scaffold-document network that represents the chemical and
+bibliometric space associated with a starting document. As an example
+of this type of analysis we constructed a network starting from a
+report describing a series of imidazonaphthalimides and their
+antitumor activity \cite{Brana:2002pr}. The network was limited to 3
+iterations (document $\rightarrow$ fragment $\rightarrow$ document
+$\rightarrow$ fragment). In addition we applied constraints on the
+fragments (e.g., must be contained in 20 to 50 molecules and exhibit a
+minimum complexity), resulting in a network with 1277 nodes (816
+document nodes and 461 fragment nodes) shown in Figure
+\ref{fig:scafodcnet}. 
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=\linewidth]{img/12477366}
+  \caption{An example of a scaffold-document network, limited to three
+    iterations. Squares represent fragments and circles represent
+    documents. Document nodes are colored based on their major topic
+    MeSH heading. The seed document is shown with a red border (top
+    left).}
+  \label{fig:scafdocnet}
+\end{figure}
+In this figure, fragment nodes are squares and document nodes are
+circles, colored by the major topic MeSH heading. (Only the 12 most
+frequent MeSH headings are considered - document nodes outside this
+set are colored grey). Though difficult to navigate as a static  
+figure, the network structure is interesting in that the immediate
+neighborhood of the seed document (red border, top left) is relatively
+small. The median number of fragments associated with a document is 1,
+(though two documents were connected to more than 100
+fragments). Conversely, the median number of documents connected to a
+fragment was 3. Fragments tend to be associated with multiple
+documents (which is expected based on the number of iterations used to
+construct the network, as well as the constraints on fragment
+characteristics) and based on the MeSH headings there is topical
+overlap between many of these documents. For example, most document
+clusters are annotated with the \emph{Chemistry} and \emph{Chemical
+  synthesis} MeSH headings. Given that the seed document has been
+annotated with \emph{Chemical synthesis} it is not surprising that
+based on the chemistry contained within it, related documents will
+also be related to synthesis. Importantly, if one considers all
+documents in ChEMBL (with valid Pubmed ID's), the two most frequent
+MeSH headings are \emph{Chemistry} and \emph{Chemical synthesis} and
+thus overrepresentation of these terms is unsurprising.
+
+This discussion has focused on the document nodes in this network. One
+could also examine the fragment nodes, in terms of their similarity to
+each other. Alternatively, one could examine the ``history'' of a
+fragment in terms of its representation in publications over
+time. Since ChEMBL also provides bioactivity data for the compounds in
+a document this network representation could be used to track the
+progression of bioactivity of a scaffold (or rather its members) over
+time.
+
 \section{R as a Platform for Network Analyses in Drug Discovery}
@@
 
+% While the \rcdk package and its dependencies provide general small
+% molecule handling functionality, a number of other packages provide
+% more specific functionality, related to small molecules. We have
+% previously mentioned DrugComboRanker \cite{huang2014drugcomboranker}
+% and MANTRA \cite{iorio2010discovery}.
+
 \section{Discussion}

### 2015-03-02 scafdoc.R

@@ -1,16 +1,77 @@
 library(igraph)
+library(RColorBrewer)
+library(Hmisc)
 
-id <- 6687479
-g <- read.graph(sprintf('4-%d.gml',id), format='gml')
+id <- 12477366
+g <- read.graph(sprintf('3-%d.xml',id), format='graphml')
 
-## filter out singletons
-#singleton.idx <- which(degree(g) > 0)
-#g <- induced.subgraph(g, vids=singleton.idx)
+## degree of doc nodes - how many fragments did we identify?
+doc.idx <- which(V(g)$klass == 'doc')
+degree(g)[doc.idx]
+frag.idx <- which(V(g)$klass == 'fragment')
+degree(g)[frag.idx]
 
-vcol <- rep('black', length(V(g)))
-vcol[ which(V(g)$klass == 'doc') ] <- 'red'
 
-pdf(sprintf("%d.pdf", id), 20,20)
+## Year based doc node colors
+year <- cut(V(g)$year, 5)
+year <- data.frame(year = levels(year))
+year.cols <- rev(brewer.pal(5, 'Blues'))[1:nrow(year)]
+year$col <- year.cols
+
+tmp <- data.frame(id=1:length(V(g)), year = cut(V(g)$year,5), klass=V(g)$klass)
+tmp <- merge(tmp, year, by='year', all.x=TRUE)
+tmp$col[ which(tmp$klass == 'fragment') ] <- 'black'
+tmp$col[ which(tmp$klass == 'doc' & is.na(tmp$col)) ] <- 'grey'
+tmp <- tmp[order(tmp$id),]
+vcol <- tmp$col
+
+## MeSH heading based doc node colors
+mesh <- data.frame(table(V(g)$mesh))
+mesh <- subset(mesh, Freq > 5 & Var1 != '' &  Var1 != 'None')
+mesh <-mesh[order(-mesh$Freq),]
+mesh.cols <- brewer.pal(nrow(mesh), 'Paired')
+mesh$col <- mesh.cols
+
+tmp <- data.frame(id=1:length(V(g)), mesh = V(g)$mesh, klass=V(g)$klass)
+tmp <- merge(tmp, mesh, by.x='mesh', by.y='Var1', all.x=TRUE)
+tmp$col[ which(tmp$klass == 'fragment') ] <- 'black'
+tmp$col[ which(tmp$klass == 'doc' & is.na(tmp$col)) ] <- 'grey'
[…]

### 2015-03-02 scafdoc.py

@@ -1,16 +1,19 @@
 import MySQLdb
 from igraph import *
 from igraph import _igraph
-import Queue
+import xml.etree.cElementTree as ET
+import Queue, urllib
+import cPickle as pickle
 
 global n
-max_round = 4
+max_round = 3
 
 db = MySQLdb.connect(host="carnot.ncats.nih.gov", # your host, usually localhost
                      user="chembl_18", # your username
                       passwd="chembl_18", # your password
                       db="chembl_18") # name o
 cur = db.cursor()
+pmid_dict = pickle.load(open('pubmed.pickle', 'rb'))
 
 def node_exists(g,name):
     try:
@@ -18,7 +21,26 @@ def node_exists(g,name):
         return True
     except:
         return False
+
+def get_pmid_meta(pmid):
+    r = pmid_dict[int(pmid)]
+    r['title'] = r['title'].encode('ascii', 'ignore')
+    return r
+
+def get_pmid_meta_old(pmid):
+    url = "http://eutils.ncbi.nlm.nih.gov/entrez/eutils/efetch.fcgi?retmode=xml&db=pubmed&id=%s" % (str(pmid))
+    page = urllib.urlopen(url).read()
+    doc = ET.fromstring(page)
+    title = doc.find("PubmedArticle/MedlineCitation/Article/ArticleTitle")
+    if title is not None: title = title.text.encode('ascii', 'ignore')
+    else: title = ""
     
+    mesh = doc.find("PubmedArticle/MedlineCitation/MeshHeadingList/MeshHeading/QualifierName[@MajorTopicYN='Y']")
+    if mesh is not None:
+        mesh = mesh.text
+    else: mesh = ""
+    return {'mesh':mesh, 'title':title}
+
 def pg(g):
     color_dict = {"doc": "red", "fragment": "black"}
     g.vs["color"] = [color_dict[klass] for klass in g.vs["klass"]]
@@ -27,7 +49,7 @@ def pg(g):
     
     plot(g, layout = layout, vertex_size=5)
 
-def get_docs_for_fragid(fid, min_year = 1997, max_year = 2015, cur = None):
+def get_docs_for_fragid(fid, min_year = 1960, max_year = 2015, cur = None):
     sql = """select distinct pubmed_id, title, abstract, year
   from ncgc_fragment_class x, ncgc_fragment_instances a, activities b, docs c
 where x.class_id = %d
@@ -38,12 +60,14 @@ and year >=  %d and year <= %d
[…]

### 2015-03-04 paper.tex
Added section on drug target networks. Updated bib entries

@@
 embodied in the ``one-target, one-drug'' approach to drug
-development. Leaving aside the subtleties of how targets are defined,
-it has become apparent that this model does not always match
-reality. Instead, it has become increasingly clear that many drugs
-(and more generally, small molecules) exhibit polypharmacological
-behavior - interacting with multiple targets, including their putative
-primary target. In many cases, the activity against the primary target
-is significantly higher than off-target activities, minimizing the
-effects of the latter. However, this is not always the case as in the
-example of terfenadine (a histamine H1 antagonist), where off target
-activity against hERG led to an increased risk of cardiac death. But
+development. Leaving aside the subtleties of how targets are defined
+\cite{Imming:2006lt}, it has become apparent that this model does not
+always match reality. Instead, it has become increasingly clear that
+many drugs (and more generally, small molecules) exhibit
+polypharmacological behavior - interacting with multiple targets,
+including their putative primary target. In many cases, the activity
+against the primary target is significantly higher than off-target
+activities, minimizing the effects of the latter. However, this is not
+always the case as in the example of terfenadine (a histamine H1
+antagonist), where off target activity against hERG
+\cite{Crumb:1995it} led to an increased risk of cardiac death. But
 more importantly, it is becoming increasingly clear that many diseases
 are not driven by dysfunction of a single a target. Examples include
-diabetes and various cancers. In parallel, even when a single target
-is the main driver for a given disease state, compensatory pathways
-can ``route around'' inhibition of the primary target, thus restoring
-the functional behavior that the primary target is responsible (i.e.,
-the system is now resistant to the original drug)
+diabetes and cancer. In parallel, even when a single target is the
+main driver for a given disease state, compensatory pathways can
+``route around'' inhibition of the primary target, thus restoring the
+functional behavior that the primary target is responsible (i.e., the
+system is now resistant to the original drug)
 \cite{Sierra:2010mz,Dienstmann:2012qr}. Finally, it should be noted
@@
 networked systems can be effective in understanding their behavior and
-then identifying ways to modulate them. The remainder of this chapter
-looks at how network methods can be applied to various topics in drug
-discovery ranging from drug-target interactions to analyzing
-structure-activity relationships.
+then identifying ways to modulate them. Importantly, network
+approaches provide the opportunity to integrate multiple entity types
+(small molecules, protein targets, genes) in a systematic fashion.The
+remainder of this chapter looks at how network methods can be applied
+to various topics in drug discovery ranging from drug-target
+interactions to analyzing structure-activity relationships.
 
@@
 \label{sec:role-networks-drug}
-Omics technologies are enabling us today to elucidate and interrogate complex relationships between disease, proteins, genetic material and small molecules at an unprecedented speed and scale.
-Interrogation of relationships reveal common mechanism of action.
-This may lead to a therapeutic candidate finding application in a seemingly unrelated disease (repositioning) that share a previously unknown causal mechanism.
-It also allows identification of drug pairs that show synergistic activity stemming from their interactions with complementary pathways.
-This type of relationship elucidation may also prevent unwanted side-effects if interrogation would reveal an unexpected link to a vital pathway that ought to remain unperturbed.
-The process of interrogation starts with experiments aimed at elucidating molecular mechanisms and pathways such as yeast two-hybrid techniques, protein and small molecule mass spectrometry, microarray and copy number variation studies.
-Experiments may also include high throughput screening of millions of chemical compounds against specific biological targets of therapeutic interest to find respective agonists and antagonists.
-These experiments result in pairwise interaction relationships from which the functional big picture biological network can then be elucidated \cite{Wu2010}.
-Algorithms that detect communities from complex networks eg.
-community structure algorithm \cite{girvan2002community} may be used.
-These network analyses algorithms allow discovery of a set of sparsely interconnected subsets of vertices where the subsets themselves are however densely connected within.
-In the biological context, this allows connecting a set of biological cycles and smaller pathways into a big picture biological metacommunity network from which inferences on cross-reactivity and new application areas of a drug candidate can be drawn.
+Omics technologies are enabling us today to elucidate and interrogate
+complex relationships between disease, proteins, genetic material and
+small molecules at an unprecedented speed and scale. Interrogation of
+relationships reveal common mechanism of action. This may lead to a
+therapeutic candidate finding application in a seemingly unrelated
+disease (repositioning) that share a previously unknown causal
+mechanism. It also allows identification of drug pairs that show
+synergistic activity stemming from their interactions with
+complementary pathways. This type of relationship elucidation may also
+prevent unwanted side-effects if interrogation would reveal an
+unexpected link to a vital pathway that ought to remain
+unperturbed. The process of interrogation starts with experiments
+aimed at elucidating molecular mechanisms and pathways such as yeast
+two-hybrid techniques, protein and small molecule mass spectrometry,
+microarray and copy number variation studies. Experiments may also
+include high throughput screening of millions of chemical compounds
+against specific biological targets of therapeutic interest to find
+respective agonists and antagonists. These experiments result in
+pairwise interaction relationships from which the functional big
+picture biological network can then be elucidated
+\cite{Wu2010}. Algorithms that detect communities from complex
+networks eg. community structure algorithm \cite{girvan2002community}
+may be used. These network analyses algorithms allow discovery of a
+set of sparsely interconnected subsets of vertices where the subsets
+themselves are however densely connected within. In the biological
+context, this allows connecting a set of biological cycles and smaller
+pathways into a big picture biological metacommunity network from
+which inferences on cross-reactivity and new application areas of a
+drug candidate can be drawn.
 
-Network based approach to mechanism elucidation of small molecules that allow identification of synergistic pairs have been implemented in tools such as the DrugComboRanker \cite{huang2014drugcomboranker}.
-The authors use a community based partitioning to create a drug metacommunity network and a disease signaling metacommunity network.
-Drug targets are then mapped onto disease networks to allow identification of synergistic pairs of drugs and complementary mechanisms of action.
-In the tool MANTRA (Mode of Action by NeTwoRk Analysis)\cite{iorio2010discovery}, authors are able to deduce mechanisms of action of new molecules and suggest repositioning of a well known molecule to a new area of application.
+Network based approach to mechanism elucidation of small molecules
+that allow identification of synergistic pairs have been implemented
+in tools such as the DrugComboRanker
+\cite{huang2014drugcomboranker}. The authors use a community based
+partitioning to create a drug metacommunity network and a disease
+signaling metacommunity network. Drug targets are then mapped onto
+disease networks to allow identification of synergistic pairs of drugs
+and complementary mechanisms of action.  In the tool MANTRA (Mode of
+Action by NeTwoRk Analysis)\cite{iorio2010discovery}, authors are able
+to deduce mechanisms of action of new molecules and suggest
+repositioning of a well known molecule to a new area of application.
 
-Network analysis has also found its application in Medicinal Chemistry and High-throughput Screening.
-Cheminformatics based chemical scaffold and topology networks capture the structure-activity relationships between compounds in screening experiments \cite{varin2011mining}.
-These help in medicinal chemistry efforts towards the identification and optimization of lead compounds for the  biological target of interest by allowing visualization of the chemical space in the backdrop of biological activity.
+Network analysis has also found its application in medicinal chemistry
+and high-throughput screening. Cheminformatics based chemical scaffold
+and topology networks capture the structure-activity relationships
+between compounds in screening experiments \cite{Varin:2011ve}. These
+help in medicinal chemistry efforts towards the identification and
+optimization of lead compounds for the biological target of interest
+by allowing visualization of the chemical space in the backdrop of
+biological activity.
 
@@
 \label{sec:drug-target-networks}
+
+Networks constructed from drugs and their protein targets were first
+described by Yildirim et al \cite{Yildirim:2007aa}, who considered
+approved and investigational drugs. Networks built using the former
+subset highlighted a number of ``follow on'' drugs (i.e., drugs
+targeting a protein that is already targeted by another drug. When
+investigational drugs were included in the analysis, the resultant
+network highlighted a more functionally diverse set of protein
+targets. Importantly, the drug-target network representation
+effectively highlights the polypharmacology exhibited by many
+drugs. They also constructed a drug-target protein network by defining
+protein targets as nodes and connecting two nodes if they are targeted
+by the same drug. But integrating this network with a protein-protein
+interaction network they were able to highlight the fact that drug
+target proteins were not as essential \cite{Jeong:2001gd} as high
+degree PPI nodes. While many drug-target networks have been
+constructed from experimentally derived or manually curated data, a
+number of methods to predicting drug-target interactions have also
+been described
+\cite{Alaimo:2013ev,Wang:2013mw,Heiskanen:2013rc}. However, it is
+still an open question as to whether such predicted networks can serve
+as the foundation for downstream analyses, though Pahikkala et al
+\cite{Pahikkala:2014gt} describe a number of guidelines to enhance the
+reliability of predicted drug-target interactions.
+
+Many workers have described analytic approaches based on drug-target
+networks. For example, the MetaCore and MetaDrug suite of tools have
+been used to perform a variety of functional analyses on small
+molecules, particularly metabolites \cite{Brennan:2009kx}.  Examples
+include pathway enrichment analyses and pathway reconstruction. The
+latter is useful in placing new or predicted metabolites in the
+context of known protein-protein and protein-compound
+interactions. Iwata et al \cite{Iwata:2013lk} described the use of a
+drug-target network to identify protein domains that are associated
+with drug side effects. Wang et al \cite{Wang:2013gn} characterized
+the topological properties of drug-target networks and noted that
+target essentiality and centrality were strongly linked to drug side
+effects and that a drugs' side effects were primarily influenced by
+the number of \emph{essential} targets rather than the number of drug
+targets. Drug-target networks have been used as input to a variety of
+predictive models. For example, a model to predict ATC codes from a
+drug-target network was described by Wang et al
+\cite{Wang:2013ye}. Tabei et al \cite{Tabei:2012tn} used a drug-target
+network a input to a classifier that was used to extract chemogenomic
+features. These features represent associations between chemical
+substructures and protein domains. The set of features that provides a
+compressed representation of the drug-target network and the authors
+proposed that they could be used to suggest new substructures for a
+given protein domain.
+
+A number of workers have used drug-target network information as the
+basis of drug repositioning methodologies. Wang et al
+\cite{Wang:2014lp} who combined drug-target and disease-target
+networks to generate repositioning hypotheses. Tan et al
+\cite{Tan:2014qb} employed drug-target networks as a component of a
+larger workflow that included chemical structure similarity, gene
+semantic similarity to identify new indications for a number of
+preexisting drugs.  Peng et al \cite{Peng:2014vy} considered a variant
+of the drug-target network concept by developing an interaction
+network of FDA approved drugs and protein pockets. Drugs were docked
+into each pocket and the docking score was used to define edges
+between the drug and protein whose pocket it was docked to.
 

### 2015-03-04 scafdoc.R

@@ -3,7 +3,7 @@ library(RColorBrewer)
 library(Hmisc)
 
 id <- 12477366
-g <- read.graph(sprintf('3-%d.xml',id), format='graphml')
+g <- read.graph(sprintf('4-%d.xml',id), format='graphml')
 
 ## degree of doc nodes - how many fragments did we identify?
 doc.idx <- which(V(g)$klass == 'doc')
@@ -52,18 +52,18 @@ vshape[ which(V(g)$klass == 'fragment') ] <- 'square'
 #vcol <- rep('black', length(V(g)))
 #vcol[ which(V(g)$klass == 'doc') ] <- 'red'
 
-pdf(sprintf("%d.pdf", id), 10,10)
+pdf(sprintf("4-%d.pdf", id), 10,10)
 
 layout(matrix(c(1,2,1,2), 2,2, byrow=TRUE), widths=c(3,1), heights=c(1,1))
 par(mar=rep(0,4))
-l = layout.fruchterman.reingold(g, niter=1000, area=50*vcount(g)^2)
+l = layout.fruchterman.reingold(g, niter=1000, area=80*vcount(g)^2)
 plot(g, layout=l, margin=0,
      vertex.size=vsize, vertex.label=NA, vertex.color=vcol, vertex.shape = vshape,
      vertex.frame.color = vframecol,
      edge.color='light grey')
 ## Legend
 par(mar=c(20,10,15,2))
-lut <- year
+lut <- mesh
 plot(rep(0, nrow(lut)), 1:nrow(lut), xlim=c(0,1),
      ylim=c(1,nrow(lut)), xaxt='n', yaxt='n',  type='n', axes=FALSE, bty='n', xlab='', ylab='')
 start.y <- 0
@@ -71,7 +71,7 @@ for (i in 1:nrow(lut)) {
   rect(0,start.y, 1,start.y+1, col=lut$col[i], border='light grey')
   start.y <- start.y+1
 }
-axis(side=2, at=seq(0.5, 4.5, by=1), labels=capitalize(as.character(lut$year)), tick=FALSE, las=2)
+axis(side=2, at=seq(0.5, 11.5, by=1), labels=capitalize(as.character(mesh$Var1)), tick=FALSE, las=2)
 dev.off()
 
 

### 2015-03-04 scafdoc.py

@@ -6,7 +6,7 @@ import Queue, urllib
 import cPickle as pickle
 
 global n
-max_round = 3
+max_round = 4
 
 db = MySQLdb.connect(host="carnot.ncats.nih.gov", # your host, usually localhost
                      user="chembl_18", # your username

### 2015-03-05 paper.tex
reworked section on networks in drug discovery

@@
 development. Leaving aside the subtleties of how targets are defined
-\cite{Imming:2006lt}, it has become apparent that this model does not
-always match reality. Instead, it has become increasingly clear that
-many drugs (and more generally, small molecules) exhibit
-polypharmacological behavior - interacting with multiple targets,
-including their putative primary target. In many cases, the activity
-against the primary target is significantly higher than off-target
-activities, minimizing the effects of the latter. However, this is not
-always the case as in the example of terfenadine (a histamine H1
-antagonist), where off target activity against hERG
-\cite{Crumb:1995it} led to an increased risk of cardiac death. But
-more importantly, it is becoming increasingly clear that many diseases
-are not driven by dysfunction of a single a target. Examples include
-diabetes and cancer. In parallel, even when a single target is the
-main driver for a given disease state, compensatory pathways can
-``route around'' inhibition of the primary target, thus restoring the
-functional behavior that the primary target is responsible (i.e., the
-system is now resistant to the original drug)
-\cite{Sierra:2010mz,Dienstmann:2012qr}. Finally, it should be noted
-that while the preceding discussion presents polypharmacological
-behavior as a negative characteristic, this is not always the
-case. Indeed, modern approaches to drug development now consider such
-behavior attractive leading to efforts to rationally design or control
-polypharmacology \cite{Gujral:2014dp,Ciceri:2014fv,Metz:2010ys}.
+\cite{Imming:2006lt}, it has become increasingly clear that many drugs
+(and more generally, small molecules) exhibit polypharmacological
+behavior - interacting with multiple targets, including their putative
+primary target. In many cases, the activity against the primary target
+is significantly higher than off-target activities, minimizing the
+effects of the latter. However, this is not always the case as in the
+example of terfenadine (a histamine H1 antagonist), where off target
+activity against hERG \cite{Crumb:1995it} led to an increased risk of
+cardiac death. Concurrently, it is becoming increasingly clear that
+many diseases are not driven by dysfunction of a single
+target. Examples include diabetes and cancer. Even when a single
+target is the main driver for a given disease state, compensatory
+pathways can ``route around'' inhibition of the primary target, thus
+restoring the functional behavior that the primary target is
+responsible (i.e., the system is now resistant to the original drug)
+\cite{Sierra:2010mz,Dienstmann:2012qr}. While the preceding discussion
+presents polypharmacological behavior as a negative characteristic,
+this is not always the case. Indeed, modern approaches to drug
+development now consider such behavior attractive, leading to efforts
+to rationally design or control polypharmacology
+\cite{Gujral:2014dp,Ciceri:2014fv,Metz:2010ys}.
 
@@
 \label{sec:role-networks-drug}
-Omics technologies are enabling us today to elucidate and interrogate
-complex relationships between disease, proteins, genetic material and
-small molecules at an unprecedented speed and scale. Interrogation of
-relationships reveal common mechanism of action. This may lead to a
-therapeutic candidate finding application in a seemingly unrelated
-disease (repositioning) that share a previously unknown causal
-mechanism. It also allows identification of drug pairs that show
-synergistic activity stemming from their interactions with
-complementary pathways. This type of relationship elucidation may also
-prevent unwanted side-effects if interrogation would reveal an
-unexpected link to a vital pathway that ought to remain
-unperturbed. The process of interrogation starts with experiments
-aimed at elucidating molecular mechanisms and pathways such as yeast
-two-hybrid techniques, protein and small molecule mass spectrometry,
-microarray and copy number variation studies. Experiments may also
-include high throughput screening of millions of chemical compounds
-against specific biological targets of therapeutic interest to find
-respective agonists and antagonists. These experiments result in
-pairwise interaction relationships from which the functional big
-picture biological network can then be elucidated
-\cite{Wu2010}. Algorithms that detect communities from complex
-networks eg. community structure algorithm \cite{girvan2002community}
-may be used. These network analyses algorithms allow discovery of a
-set of sparsely interconnected subsets of vertices where the subsets
-themselves are however densely connected within. In the biological
-context, this allows connecting a set of biological cycles and smaller
-pathways into a big picture biological metacommunity network from
-which inferences on cross-reactivity and new application areas of a
-drug candidate can be drawn.
 
-Network based approach to mechanism elucidation of small molecules
-that allow identification of synergistic pairs have been implemented
-in tools such as the DrugComboRanker
-\cite{huang2014drugcomboranker}. The authors use a community based
-partitioning to create a drug metacommunity network and a disease
-signaling metacommunity network. Drug targets are then mapped onto
-disease networks to allow identification of synergistic pairs of drugs
-and complementary mechanisms of action.  In the tool MANTRA (Mode of
-Action by NeTwoRk Analysis)\cite{iorio2010discovery}, authors are able
-to deduce mechanisms of action of new molecules and suggest
-repositioning of a well known molecule to a new area of application.
+Computational approaches in drug discovery, such as docking and
+statistical QSAR, tend to focus on interactions between small
+molecules and protein targets. Network representations and models play
+a variety of roles in the \emph{in silico} drug discovery pipeline
+ranging from visualization of heteregeneous datasets
+\cite{Huan:2010cy,Kuhn:2010zr} to developing predictive models of drug
+action \cite{Vina:2009oq,Folger:2011tx}. A necessary requirement for
+network based analyses is the availability of network (or interaction)
+data. This can come in many forms. Examples include protein-protein
+interaction (PPI) databases such as BioGRID
+\cite{Chatr-Aryamontri:2015yf} and STRING \cite{Franceschini:2013qa}
+and drug-target databases such as DrugBank
+\cite{Wishart:2008aa}. Bioactivity databases such as Pubchem and
+ChEMBL \cite{Gaulton:2012jl} also provide useful information on
+drug-target relationships, based on measured binding affinities.  In
+addition to small molecules and proteins, databases such as PharmGKB
+\cite{Thorn:2005aa} and MetaCore (a commercial product from Thompson
+Reuters) integrate small molecules, protein and gene targets and the
+published literature, providing a rich source of relationships between
+these entities.
 
-Network analysis has also found its application in medicinal chemistry
-and high-throughput screening. Cheminformatics based chemical scaffold
-and topology networks capture the structure-activity relationships
-between compounds in screening experiments \cite{Varin:2011ve}. These
-help in medicinal chemistry efforts towards the identification and
-optimization of lead compounds for the biological target of interest
-by allowing visualization of the chemical space in the backdrop of
-biological activity.
+Using these databases, network methods can be applied to various
+stages of the drug discovery pipeline. For example, target
+identification can benefit from network analyses by characterizing the
+role of a given target in the context of the larger interaction and
+regulatory network. Thus targeting proteins that have a high degree of
+connectedness may lead to significant side effects in other parts of
+the system [REF] and can be the source of observed side effects. On
+these lines, approaches to characterize toxicity have employed network
+analyses \cite{Kim:2014xs,Zhang:2014hj}. While some approaches have
+considered side effects [REF] others have considered the problem at
+the genomic level by characterizing gene networks that are perturbed
+by small molecules. Campaigns focused on combination therapies can
+also benefit from a network representation and network inference
+approaches \cite{Huang:2014by,Tang:2013lr} to identify target
+combinations whose simultaneous inhibition leads to synergistic
+effects. Mechanism of action studies can benefit from network analyses
+by integrating multiple data types. An example of this was described
+by Young et al \cite{Young:2008aa} that integrated high content
+screening and ligand-target prediction models to construct network
+representations of MoA hypotheses. Network approaches have also been
+proposed for SAR analysis \cite{Iyer:2011ij,Wawer:2008aa}, usually at
+the lead optimization stage of the discovery pipeline. Drug
+repositioning efforts \cite{Dudley:2011dn,Gottlieb:2011rx} benefit
+from network analyses such as community detection algorithms
+\cite{girvan2002community}.
 
@@
 between small molecules, protein or gene targets and diseases allow
-one to develop network representations which can then be visualized
-and quantified to support integrative analyses of multiple data
-types. In the following subsections we highlight applications of this
-approach and where possible provide examples of R code to generate and
-analyze such networks.
+one to develop network representations of these entities which can
+then be visualized and quantified to support integrative analyses of
+multiple data types. In the following subsections we highlight
+applications of this approach and where possible provide examples of R
+code to generate and analyze such networks.
 

### 2015-03-06 code.R
regenerated similarity network graphs

@@ -72,7 +72,8 @@ vcols <- unlist(sapply(V(g2)$panther, function(x) col21[x]))
 pdf('img/sim-network-connected.pdf', 9,7)
 layout(matrix(c(1,2,1,2), 2,2, byrow=TRUE), widths=c(3,1), heights=c(1,1))
 par(mar=c(0,0,0,0))
-plot(g2, layout=layout.fruchterman.reingold,
+l <- layout.fruchterman.reingold(g2, niter=1000, area=80*vcount(g2)^2)
+plot(g2, layout=l,
      margin=0,
      vertex.label=NA, vertex.size=2.5, vertex.color=vcols)
 par(mar=c(0,0,0,0.5))

### 2015-03-06 paper.tex
first draft done

@@
 statistical QSAR, tend to focus on interactions between small
-molecules and protein targets. Network representations and models play
-a variety of roles in the \emph{in silico} drug discovery pipeline
-ranging from visualization of heteregeneous datasets
-\cite{Huan:2010cy,Kuhn:2010zr} to developing predictive models of drug
-action \cite{Vina:2009oq,Folger:2011tx}. A necessary requirement for
-network based analyses is the availability of network (or interaction)
-data. This can come in many forms. Examples include protein-protein
-interaction (PPI) databases such as BioGRID
+molecules and protein targets. By definition they do not consider the
+surrounding functional environment of interacting proteins and related
+pathways. Network representations on the other hand are capable of
+capturing the functional context of individual molecular entities and
+network models play a variety of roles in the \emph{in silico} drug  
+discovery pipeline ranging from visualization of heteregeneous
+datasets \cite{Huan:2010cy,Kuhn:2010zr} to developing predictive
+models of drug action \cite{Vina:2009oq,Folger:2011tx}. A necessary
+requirement for network based analyses is the availability of network
+(or interaction) data. This can come in many forms. Examples include
+protein-protein interaction (PPI) databases such as BioGRID
 \cite{Chatr-Aryamontri:2015yf} and STRING \cite{Franceschini:2013qa}
@@
 regulatory network. Thus targeting proteins that have a high degree of
-connectedness may lead to significant side effects in other parts of
-the system [REF] and can be the source of observed side effects. On
-these lines, approaches to characterize toxicity have employed network
-analyses \cite{Kim:2014xs,Zhang:2014hj}. While some approaches have
-considered side effects [REF] others have considered the problem at
-the genomic level by characterizing gene networks that are perturbed
-by small molecules. Campaigns focused on combination therapies can
-also benefit from a network representation and network inference
-approaches \cite{Huang:2014by,Tang:2013lr} to identify target
-combinations whose simultaneous inhibition leads to synergistic
-effects. Mechanism of action studies can benefit from network analyses
-by integrating multiple data types. An example of this was described
-by Young et al \cite{Young:2008aa} that integrated high content
-screening and ligand-target prediction models to construct network
-representations of MoA hypotheses. Network approaches have also been
-proposed for SAR analysis \cite{Iyer:2011ij,Wawer:2008aa}, usually at
-the lead optimization stage of the discovery pipeline. Drug
-repositioning efforts \cite{Dudley:2011dn,Gottlieb:2011rx} benefit
-from network analyses such as community detection algorithms
+connectedness (as measured by essentiality and centrality) may lead to
+significant side effects in other parts of the system and can be the
+source of observed side effects \cite{Wang:2013gn}. On these lines,
+approaches to characterize toxicity have employed network analyses
+\cite{Kim:2014xs,Zhang:2014hj}. While some approaches have considered
+side effects [REF] others have considered the problem at the genomic
+level by characterizing gene networks that are perturbed by small
+molecules. Campaigns focused on combination therapies can also benefit
+from a network representation and network inference approaches
+\cite{Huang:2014by,Tang:2013lr} to identify target combinations whose
+simultaneous inhibition leads to synergistic effects. Mechanism of
+action studies can benefit from network analyses by integrating
+multiple data types. An example of this was described by Young et al
+\cite{Young:2008aa} that integrated high content screening and
+ligand-target prediction models to construct network representations
+of MoA hypotheses. Network approaches have also been proposed for SAR
+analysis \cite{Iyer:2011ij,Wawer:2008aa}, usually at the lead
+optimization stage of the discovery pipeline. Drug repositioning
+efforts \cite{Dudley:2011dn,Gottlieb:2011rx} benefit from network
+analyses such as community detection algorithms
 \cite{girvan2002community}.
+
+While drugs can range from small molecules to antibodies and from pure
+compounds to mixtures, the following sections focus on how network
+approaches have been used to characterize and model the interactions
+of small molecules with biological systems. We group such interactions
+based on the nature of the interacting entities and provide a brief
+overview of recent work on how network representations have been
+employed to analyze them.
 
@@
 
-% While the \rcdk package and its dependencies provide general small
-% molecule handling functionality, a number of other packages provide
-% more specific functionality, related to small molecules. We have
-% previously mentioned DrugComboRanker \cite{huang2014drugcomboranker}
-% and MANTRA \cite{iorio2010discovery}.
-
 \section{Discussion}
@@
 
-In this chapter we have attempted to provide an overview of how
-networks have been employed to analyze and visualize chemical
-data. The fact that chemical structures themselves are graphs allows
+In this chapter we have summarized how
+networks have been employed to analyze and visualize small molecule
+data (structures, interactions, activity) in the context of protein
+and gene targets. The fact that chemical structures themselves are graphs allows
 one to apply a number of techniques from graph theory to characterize
@@
 rather than using the network structure to draw insight into the
-underlying chemical of biological phenomena. This is not to say that
+underlying chemical or biological phenomena. This is not to say that
 network visualizations of chemical data are lacking. In fact, such
 visualizations allow one to encode multiple properties of objects
-(i.e., nodes) and relationships (i.e., edges) in a single visual form.
+(i.e., nodes) and relationships (i.e., edges) in a single visual
+form. However, in many cases, studies tend to report structural
+characterizations (average vertex degree, centrality values and so on)
+but do not make a persuasive link between these and the underlying
+chemistry and biology.
 
@@
 
+In conclusion, network representations of molecular entities and their
+interactions are a powerful tool that captures (possibly) large and
+heterogenous datasets. Network approaches have been employed at
+various stages of the drug discovery pipeline ranging from target
+identification to SAR analysis. While one could argue that many
+results from network approaches could be achieved by simpler tabular
+queries, it is evident that in a number of cases, the integrative
+nature of network methods provides a novel view of the data and opens
+up alternative avenues of investigation.
+
+\section{Acknowledgements}
+\label{sec:acknowledgements}
+
+We would like to thank Dac-Trung Nguyen and Tyler Peryea for fruitful
+discussions related scaffold and scaffold-document networks.
+
 \bibliography{paper}

### 2015-03-09 paper.tex
Update paper.tex

@@
 \title{Exploring the Role of Small Molecules in Biological Systems Using Network Approaches}
-\author{Rajarshi Guha and Saurav Das}
+\author{Rajarshi Guha and Sourav Das}
 
@@
 between the drug and protein whose pocket it was docked to.
+
+Other tools that utilize drug-target networks include the DrugComboRanker \cite{huang2014drugcomboranker} where the authors use a community based partitioning to create a drug metacommunity network and a disease signaling metacommunity network.
+Drug targets are then mapped onto disease networks to allow identification of synergistic pairs of drugs and complementary mechanisms of action.
+In the tool MANTRA (Mode of Action by NeTwoRk Analysis)\cite{iorio2010discovery}, authors apply the concept of network communities to deduce mechanism of action of new molecules and suggest repositioning of a well known molecule to a new area of therapeutic application.
+A recently developed package in R called the TIMMA-R package by He et al.
+(2015) combines binary drug-target data and drug-sensitivity profiles to predict synergistic pairs of anti-cancer drugs.
+The package is freely available from the CRAN repository.
 

### 2015-03-20 paper.tex
Update paper.tex
Updated section 1 with a description of the chapter

--- 2015-03-06 paper.tex
@@
 development. Leaving aside the subtleties of how targets are defined
-\cite{Imming:2006lt}, it has become increasingly clear that many drugs
-(and more generally, small molecules) exhibit polypharmacological
-behavior - interacting with multiple targets, including their putative
-primary target. In many cases, the activity against the primary target
-is significantly higher than off-target activities, minimizing the
-effects of the latter. However, this is not always the case as in the
-example of terfenadine (a histamine H1 antagonist), where off target
-activity against hERG \cite{Crumb:1995it} led to an increased risk of
-cardiac death. Concurrently, it is becoming increasingly clear that
-many diseases are not driven by dysfunction of a single
-target. Examples include diabetes and cancer. Even when a single
-target is the main driver for a given disease state, compensatory
-pathways can ``route around'' inhibition of the primary target, thus
-restoring the functional behavior that the primary target is
-responsible (i.e., the system is now resistant to the original drug)
-\cite{Sierra:2010mz,Dienstmann:2012qr}. While the preceding discussion
-presents polypharmacological behavior as a negative characteristic,
-this is not always the case. Indeed, modern approaches to drug
-development now consider such behavior attractive, leading to efforts
-to rationally design or control polypharmacology
+\cite{Imming:2006lt}, it has become increasingly clear that many small
+molecules exhibit polypharmacological behavior - interacting with
+multiple targets, including their putative primary target. In many
+cases, the activity against the primary target is significantly higher
+than off-target activities, minimizing the effects of the
+latter. However, this is not always the case as in the example of
+terfenadine (a histamine H1 antagonist), where off target activity
+against hERG \cite{Crumb:1995it} led to an increased risk of cardiac
+death. Concurrently, it is becoming increasingly clear that many
+diseases are not driven by dysfunction of a single target. Examples
+include diabetes and cancer. Even when a single target is the main
+driver for a given disease state, compensatory pathways can ``route
+around'' inhibition of the primary target, thus restoring the
+functional behavior that the primary target is responsible (i.e., the
+system is now resistant to the original drug)
+\cite{Sierra:2010mz,Dienstmann:2012qr}. While this discussion presents
+polypharmacological behavior as a negative characteristic, this is not
+always the case. Indeed, modern approaches to drug development now
+consider such behavior attractive, leading to efforts to rationally
+design or control polypharmacology
 \cite{Gujral:2014dp,Ciceri:2014fv,Metz:2010ys}.
@@
 approaches provide the opportunity to integrate multiple entity types
-(small molecules, protein targets, genes) in a systematic fashion.The
-remainder of this chapter looks at how network methods can be applied
-to various topics in drug discovery ranging from drug-target
-interactions to analyzing structure-activity relationships.
+(small molecules, protein targets, genes) in a systematic fashion.
+
+The remainder of this chapter looks at how network methods can be
+applied to various topics in drug discovery ranging from drug-target
+interactions to analyzing structure-activity relationships. In
+addition to reviewing methodologies, we also highlight implementations
+of these methods and related tools that are available for network
+analyses in the R environment. We conclude the chapter with a
+discussion of working with chemical structure data in the R
+environment and how such structures can be employed in a network
+analysis.
 
@@
 
-Using these databases, network methods can be applied to various
+Using these databases, network methods can be applied at various
 stages of the drug discovery pipeline. For example, target
@@
 analysis \cite{Iyer:2011ij,Wawer:2008aa}, usually at the lead
-optimization stage of the discovery pipeline. Drug repositioning
-efforts \cite{Dudley:2011dn,Gottlieb:2011rx} benefit from network
-analyses such as community detection algorithms
-\cite{girvan2002community}.
+optimization stage. Drug repositioning efforts
+\cite{Dudley:2011dn,Gottlieb:2011rx} benefit from network analyses
+such as community detection algorithms \cite{girvan2002community}.
 
@@
 derive invariants (also known as topological descriptors
-\cite{Guha:2012vn}). However, small molecules interact with protein
-targets and the targets themselves interact with each other. Such
-explicit interactions lend themselves naturally to network
-representations. More generally, observed or computed relationships
-between small molecules, protein or gene targets and diseases allow
-one to develop network representations of these entities which can
-then be visualized and quantified to support integrative analyses of
-multiple data types. In the following subsections we highlight
-applications of this approach and where possible provide examples of R
-code to generate and analyze such networks.
+\cite{Guha:2012vn}). In this work we focus on small molecule
+interactions with protein targets and the targets themselves interact
+with each other. Such explicit interactions lend themselves naturally
+to network representations. More generally, observed or computed
+relationships between small molecules, protein or gene targets and
+diseases allow one to develop network representations of these
+entities which can then be visualized and quantified to support
+integrative analyses of multiple data types. In the following
+subsections we highlight applications of this approach and where
+possible provide examples of R code to generate and analyze such
+networks.
 
@@
 \cite{Wang:2013ye}. Tabei et al \cite{Tabei:2012tn} used a drug-target
-network a input to a classifier that was used to extract chemogenomic
+network as input to a classifier that was used to extract chemogenomic
 features. These features represent associations between chemical
@@
 between the drug and protein whose pocket it was docked to.
+
+A variety of algorithms related to repositioning and activity
+predictions, based on drug-target networks have been implemented in
+R. Ma and Zhao \cite{Ma:2012qy} implemented the FacPad method that
+utilizes a Bayesian sparse factor model to identify pathways that are
+perturbed upon drug treatments. The DrugComboRanker package
+\cite{Huang:2014by} employed a community based partitioning to create
+a drug meta-community network and a disease signaling meta-community
+network. Drug targets were then mapped onto disease networks to allow
+identification of synergistic pairs of drugs and complementary
+mechanisms of action. In the tool called MANTRA (Mode of Action by
+NeTwoRk Analysis) \cite{iorio2010discovery}, authors used network
+communities to deduce mechanism of action of new molecules and suggest
+repositioning of a well known molecule to a new therapeutic
+arean. Finally the TIMMA-R package \cite{He:2015ly} combines binary
+drug-target data and drug-sensitivity profiles to predict synergistic
+pairs of anti-cancer drugs.
 
@@
 We would like to thank Dac-Trung Nguyen and Tyler Peryea for fruitful
-discussions related scaffold and scaffold-document networks.
+discussions related to scaffold and scaffold-document networks.
 

### 2015-03-20 scafdoc.R

@@ -3,6 +3,7 @@ library(RColorBrewer)
 library(Hmisc)
 
 id <- 12477366
+##id <- 21568322
 g <- read.graph(sprintf('4-%d.xml',id), format='graphml')
 
 ## degree of doc nodes - how many fragments did we identify?
@@ -39,6 +40,20 @@ tmp$col[ which(tmp$klass == 'doc' & is.na(tmp$col)) ] <- 'grey'
 tmp <- tmp[order(tmp$id),]
 vcol <- tmp$col
 
+## Target classification headings
+tc <- data.frame(table(V(g)$l2))
+tc <- tc[order(tc$Freq),]
+tc <- subset(tc, Freq > 3)
+tc.cols <- brewer.pal(nrow(tc), 'Paired')
+tc$col <- tc.cols
+
+tmp <- data.frame(id=1:vcount(g), tc = V(g)$l2, klass=V(g)$klass)
+tmp <- merge(tmp, tc, by.x='tc', by.y='Var1', all.x=TRUE)
+tmp$col[ which(tmp$klass == 'fragment') ] <- 'black'
+tmp$col[ which(tmp$klass == 'doc' & is.na(tmp$col)) ] <- 'grey'
+tmp <- tmp[order(tmp$id),]
+vcol <- tmp$col
+
 vframecol <- rep('black', length(V(g)))
 vframecol[1] <- 'red'
 
@@ -63,7 +78,7 @@ plot(g, layout=l, margin=0,
      edge.color='light grey')
 ## Legend
 par(mar=c(20,10,15,2))
-lut <- mesh
+lut <-  tc
 plot(rep(0, nrow(lut)), 1:nrow(lut), xlim=c(0,1),
      ylim=c(1,nrow(lut)), xaxt='n', yaxt='n',  type='n', axes=FALSE, bty='n', xlab='', ylab='')
 start.y <- 0
@@ -71,7 +86,7 @@ for (i in 1:nrow(lut)) {
   rect(0,start.y, 1,start.y+1, col=lut$col[i], border='light grey')
   start.y <- start.y+1
 }
-axis(side=2, at=seq(0.5, 11.5, by=1), labels=capitalize(as.character(mesh$Var1)), tick=FALSE, las=2)
+axis(side=2, at=seq(0.5, nrow(lut)-0.5, by=1), labels=capitalize(as.character(lut$Var1)), tick=FALSE, las=2)
 dev.off()
 
 

### 2015-03-20 scafdoc.py

@@ -4,7 +4,7 @@ from igraph import _igraph
 import xml.etree.cElementTree as ET
 import Queue, urllib
 import cPickle as pickle
-
+from itertools import groupby
 global n
 max_round = 4
 
@@ -23,9 +23,12 @@ def node_exists(g,name):
         return False
 
 def get_pmid_meta(pmid):
-    r = pmid_dict[int(pmid)]
-    r['title'] = r['title'].encode('ascii', 'ignore')
-    return r
+    try:
+        r = pmid_dict[int(pmid)]
+        r['title'] = r['title'].encode('ascii', 'ignore')
+        return r
+    except:
+        return {'title':None, 'mesh':None}
 
 def get_pmid_meta_old(pmid):
     url = "http://eutils.ncbi.nlm.nih.gov/entrez/eutils/efetch.fcgi?retmode=xml&db=pubmed&id=%s" % (str(pmid))
@@ -50,21 +53,41 @@ def pg(g):
     plot(g, layout = layout, vertex_size=5)
 
 def get_docs_for_fragid(fid, min_year = 1960, max_year = 2015, cur = None):
-    sql = """select distinct pubmed_id, title, abstract, year
-  from ncgc_fragment_class x, ncgc_fragment_instances a, activities b, docs c
+    sql = """select distinct pubmed_id, title, abstract, year, l1, l2, l3, l4, l5
+  from ncgc_fragment_class x, ncgc_fragment_instances a, activities b, docs c, assays ass,
+    target_dictionary td,
+	target_components tc,
+    component_sequences cs,
+    component_class cc,
+    protein_family_classification pfc
 where x.class_id = %d
 and x.class_id = a.class_id
 and a.molregno = b.molregno
 and c.doc_id = b.doc_id
 and year >=  %d and year <= %d
-and pubmed_id is not null""" % (fid, min_year, max_year)
+and pubmed_id is not null
+ and c.doc_id = ass.doc_id
+        and ass.tid = td.tid
+		and td.tax_id = 9606
+		and td.tid = tc.tid
+		AND tc.component_id = cs.component_id
+        AND cc.component_id = cs.component_id
+        AND pfc.protein_class_id = cc.protein_class_id""" % (fid, min_year, max_year)
 
     cur.execute(sql)
     rows = cur.fetchall()
-    rows = [ [str(x[0]), x[1], x[2], x[3]] for x in rows]
-    return rows
[…]

### 2015-03-22 paper.tex
updated with increased emphasis on R packages for network analysis

@@
 \cite{Kim:2014xs,Zhang:2014hj}. While some approaches have considered
-side effects [REF] others have considered the problem at the genomic
-level by characterizing gene networks that are perturbed by small
-molecules. Campaigns focused on combination therapies can also benefit
-from a network representation and network inference approaches
-\cite{Huang:2014by,Tang:2013lr} to identify target combinations whose
-simultaneous inhibition leads to synergistic effects. Mechanism of
-action studies can benefit from network analyses by integrating
-multiple data types. An example of this was described by Young et al
-\cite{Young:2008aa} that integrated high content screening and
-ligand-target prediction models to construct network representations
-of MoA hypotheses. Network approaches have also been proposed for SAR
-analysis \cite{Iyer:2011ij,Wawer:2008aa}, usually at the lead
-optimization stage. Drug repositioning efforts
+side effects \cite{Lounkine:2012yf} others have considered the problem
+at the genomic level by characterizing gene networks that are
+perturbed by small molecules. Campaigns focused on combination
+therapies can also benefit from a network representation and network
+inference approaches \cite{Huang:2014by,Tang:2013lr} to identify
+target combinations whose simultaneous inhibition leads to synergistic
+effects. Mechanism of action studies can benefit from network analyses
+by integrating multiple data types. An example of this was described
+by Young et al \cite{Young:2008aa} that integrated high content
+screening and ligand-target prediction models to construct network
+representations of MoA hypotheses. Network approaches have also been
+proposed for SAR analysis \cite{Iyer:2011ij,Wawer:2008aa}, usually at
+the lead optimization stage. Drug repositioning efforts
 \cite{Dudley:2011dn,Gottlieb:2011rx} benefit from network analyses
@@
 employed to analyze them.
+
+\section{R for Network Analyses}
+\label{sec:r-network-analyses}
+
+While this chapter focuses on the use of network methodologies in
+various drug discovery related problems, we also focus on the use of R
+for these analyses. Section \ref{sec:r-as-platform} discusses R
+packages oriented towards handling molecular data types such as
+chemical structures and genomic data. In this section we briefly
+review the packages that are available for handling and analyzing
+network datasets. The \texttt{igraph} \cite{Csardi:2006lp} and
+\texttt{network} \cite{Butts:2008tp} packages represent comprehensive
+tools to construct, manipulate and visualize network datasets within
+the R environment. The \texttt{graph} package \cite{Gentleman:1wt}
+also aims to support general graph manipulation, but is somewhat
+limited in scope compared to \texttt{igraph} and \texttt{network}
+(both of which share some aspects of syntax and interfaces). These
+three packages are the basis of many other network analysis and visualization
+packages. An example is \texttt{statnet} \cite{Handcock:2008il} which
+employs the \texttt{network} package and focuses on the use of
+exponential random graph models \cite{Robins:2007cr} to support
+statistical inference and simulation.
+
+While many network analysis packages in R focus on social science
+applications, there are a plethora of R packages that focus on
+biological networks. For example, the WCGNA package
+\cite{Langfelder:2008cj} implements methods to construct gene networks
+based on co-expression data. The Corbi package \cite{Huang:2013sy}
+implements the CNetQ algorithm for network alignment, a task that is
+relevant for predicting new protein-protein interactions. Given the
+integrative nature of many biological problems that require the use of
+multiple data types (e.g., genomic, proteomic, small molecule),
+network approaches are commonly applied to investigating them. A
+number of R packages support such integrative analyses - BioNet
+\cite{Beisser:2010xp} is an R package that supports a variety of
+integrative analyses, with a particular focus on identifying
+functional modules in $k$-partite networks. NetWeAvers
+\cite{McClellan:2013zf} is similar in nature, though it has
+specialized support for mass spectrometry data. A number of packages
+support statistical operations on graphs. For example,
+\texttt{DEGraph} implements a method to test for differential
+expression of gene networks \cite{Jacob:2012yg}. The \texttt{SANTA}
+package \cite{Cornish:2014vs} implements methods to test the
+association between a network and a phenotype. Several packages are
+also available for enrichment analyses including \texttt{FGNet},
+\texttt{miRNApath} and \texttt{viper}.
 
@@
 opportunities for repurposing drugs based on novel disease-target or
-disease-pathway associations. Another view provided by such analysis
+disease-pathway associations. For example, Pacini et al
+\cite{Pacini:2013kl} describe DvD, a pipeline comprised of R and
+Cytoscape \cite{Shannon:2003aa} to identify repurposing opportunities
+using gene expression profiles. Another view provided by such analysis
 is the druggability of a disease; that is, based on underlying targets
@@
 targets for diseases or links between diseases, but instead attempt to
-elucidate the mechanisms underlying condition. As an example of this is
-the paradoxical effects of MEK inhibition and PI3K/AKT inhibition and
-the effects on cancer resistance \cite{Sos:2009gc}. Together with knowledge
-of the underlying network connectivity, these approaches allow one to
-design combination treatments - multiple drugs that affect multiple
-targets and pathways. This approach was exemplified by Lehar et al
-\cite{Lehar:2009gu} who proposed that drug combination response
-surfaces were functionally related to the underlying connectivity of
-the drug targets.
+elucidate the mechanisms underlying condition. As an example of this
+is the paradoxical effects of MEK inhibition and PI3K/AKT inhibition
+and the effects on cancer resistance \cite{Sos:2009gc}. Together with
+knowledge of the underlying network connectivity, these approaches
+allow one to design combination treatments - multiple drugs that
+affect multiple targets and pathways. This approach was exemplified by
+Lehar et al \cite{Lehar:2009gu} who proposed that drug combination
+response surfaces were functionally related to the underlying
+connectivity of the drug targets.
 
@@
 document nodes and 461 fragment nodes) shown in Figure
-\ref{fig:scafodcnet}. 
+\ref{fig:scafdocnet}.
+
 \begin{figure}[h]
@@
 methods to be manipulate them have been implemented in a variety of
-software tools. A comprehensive computational drug discovery platform
-will include a number of these tools, as it is rare that a single tool
-addresses all aspects.
+software tools. These range from software libraries such as the CDK
+\cite{Steinbeck:2006aa} or RDKit \cite{RDKit:2008aa} for manipulating
+chemical structures to pipeline tools such as KNIME
+\cite{KNIME:2008aa} or PipelinePilot that can be used to integrate
+disparate tools and data sources using a graphical user interface.  A
+comprehensive computational drug discovery infrastructure will include
+a number of these tools, as it is rare that a single tool addresses
+all aspects.
 
-In this sense, the R environment presents a robust platform for
+Statistical modeling and machine learning methods have become an
+integral part of computational drug discovery. Many drug discovery
+tools and platforms provide access to statistical methods. In some
+cases such as in RDKit, these are implemented from scratch, where as
+in others such as PipelinePilot, external statistical tools such as R
+and JMP are integrated.
+
+If one is primarily interested in the statistical analyses in a drug
+discovery setting, the R environment presents a robust platform for
 computational drug discovery, especially the stages up to and
-including lead optimization. The Bioconductor set of packages provide
-a plethora of tools to analyze a variety of ``omics'' data. Examples
-include packages that support gene expression (microarray or RNAseq)
-analyses, tools for differential expression analysis, gene and pathway
-enrichment analysis and so on.
+including lead optimization.  It is well known that many cutting edge
+statistical techniques are first implemented in R and this is
+increasingly becoming the case for much of bioinformatics. The
+Bioconductor set of packages provides a plethora of tools to analyze a
+variety of ``omics'' data. Examples include packages that support gene
+expression (microarray or RNAseq) analyses, tools for differential
+expression analysis, gene and pathway enrichment analysis and so on.
 
-In contrast, support for small molecule data is limited. Currently,
+However, support for small molecule data is limited. A number of
+packages focus on specific cheminformatics related
+tasks. Examples include the \texttt{Rcpi} package that computes
+protein ligand descriptors \cite{Cao:2015uq} , the \texttt{QuACN} package that
+computes a number of graph invariants \cite{Mueller:2011rz} from a
+molecular graph and the \texttt{RMol} package to parse MOL files
+\cite{Grabner:2012rr}. However, in many scenarios it is useful to
+be able to manipulate chemical structures at a lower level. Thus one
+requires access to core cheminformatics functionality. Currently,
 there are two packages that focus on representing and manipulating
@@
 implemented in an underlying Java library (JOELib for ChemmineR and
-the CDK for \rcdk). In this section we focus on the use of \rcdk as a
-means to manipulate and analyze chemical structure data.
+the CDK for \rcdk). These packages provide a low level interface to
+cheminformatics libraries, allowing the R user to manipulate data
+structures representing atoms, bonds and molecules as well as perform
+core cheminformatics tasks such as atom typing and aromaticity
+perception. But in addition, higher level features are also made
+available including descriptor calculation, substructure searching,
+fragmentation and 2D structure depictions. In this section we focus on
+the use of \rcdk a means to manipulate and analyze chemical structure
+data.
 
@@
 \end{lstlisting}
-In this code snippet, we compute a pairwise similarity matrix that we
-will eventually use as an adjacency matrix to construct the similarity
-network. We only keep those similarity values greater than 0.6 so as
-to focus only on similar molecules.
+In this snippet, we computed hashed path fingerprints as implemented
+in the CDK, using the default values of length (1024) and depth
+(6). The CDK supports multiple fingerprint types including
+substructural keys (MACCS, Pubchem) and circular fingerprints (ECFP)
+and we can easily switch to another one by changing the \texttt{type}
+argument. Once the fingerprints are computed we compute a pairwise
+similarity matrix that we will eventually use as an adjacency matrix
+to construct the similarity network. We only keep those similarity
+values greater than 0.6 so as to focus only on similar molecules.
 
@@
 
-In this chapter we have summarized how
-networks have been employed to analyze and visualize small molecule
-data (structures, interactions, activity) in the context of protein
-and gene targets. The fact that chemical structures themselves are graphs allows
-one to apply a number of techniques from graph theory to characterize
-chemical structure (such as the many graph invariants). However, with
-the advent of large collections of chemical structures (from databases
-such as Pubchem or high throughput screening programs) along with
-associated data such as bioactivity and target annotations, a number
-of workers have attempted to analyze the properties of such
-collections using network techniques.
+In this chapter we have summarized how networks have been employed to
+analyze and visualize small molecule data (structures, interactions,
+activity) in the context of protein and gene targets. The fact that
+chemical structures themselves are graphs allows one to apply a number
+of techniques from graph theory to characterize chemical structure
+(such as the many graph invariants). However, with the advent of large
+collections of chemical structures (from databases such as Pubchem or
+high throughput screening programs) along with associated data such as
+bioactivity and target annotations, a number of workers have attempted
+to analyze the properties of such collections using network
+techniques.
 

### 2015-03-22 paper.tex
Updated some formatting

@@
 applications, there are a plethora of R packages that focus on
-biological networks. For example, the WCGNA package
+biological networks. For example, the \texttt{WCGNA} package
 \cite{Langfelder:2008cj} implements methods to construct gene networks
-based on co-expression data. The Corbi package \cite{Huang:2013sy}
+based on co-expression data. The \texttt{Corbi} package \cite{Huang:2013sy}
 implements the CNetQ algorithm for network alignment, a task that is
@@
 network approaches are commonly applied to investigating them. A
-number of R packages support such integrative analyses - BioNet
+number of R packages support such integrative analyses - \texttt{BioNet}
 \cite{Beisser:2010xp} is an R package that supports a variety of
 integrative analyses, with a particular focus on identifying
-functional modules in $k$-partite networks. NetWeAvers
+functional modules in $k$-partite networks. \texttt{NetWeAvers}
 \cite{McClellan:2013zf} is similar in nature, though it has
@@
 predictions, based on drug-target networks have been implemented in
-R. Ma and Zhao \cite{Ma:2012qy} implemented the FacPad method that
-utilizes a Bayesian sparse factor model to identify pathways that are
-perturbed upon drug treatments. The DrugComboRanker package
-\cite{Huang:2014by} employed a community based partitioning to create
-a drug meta-community network and a disease signaling meta-community
-network. Drug targets were then mapped onto disease networks to allow
-identification of synergistic pairs of drugs and complementary
-mechanisms of action. In the tool called MANTRA (Mode of Action by
-NeTwoRk Analysis) \cite{iorio2010discovery}, authors used network
-communities to deduce mechanism of action of new molecules and suggest
-repositioning of a well known molecule to a new therapeutic
-arean. Finally the TIMMA-R package \cite{He:2015ly} combines binary
-drug-target data and drug-sensitivity profiles to predict synergistic
-pairs of anti-cancer drugs.
+R. Ma and Zhao \cite{Ma:2012qy} implemented the \texttt{FacPad} method
+that utilizes a Bayesian sparse factor model to identify pathways that
+are perturbed upon drug treatments. The \texttt{DrugComboRanker}
+package \cite{Huang:2014by} employed a community based partitioning to
+create a drug meta-community network and a disease signaling
+meta-community network. Drug targets were then mapped onto disease
+networks to allow identification of synergistic pairs of drugs and
+complementary mechanisms of action. In the tool called \texttt{MANTRA}
+(Mode of Action by NeTwoRk Analysis) \cite{iorio2010discovery},
+authors used network communities to deduce mechanism of action of new
+molecules and suggest repositioning of a well known molecule to a new
+therapeutic arean. Finally the \texttt{TIMMA-R} package
+\cite{He:2015ly} combines binary drug-target data and drug-sensitivity
+profiles to predict synergistic pairs of anti-cancer drugs.
 

### 2015-03-22 paper.tex
Updated et al

@@
 by integrating multiple data types. An example of this was described
-by Young et al \cite{Young:2008aa} that integrated high content
+by Young et al. \cite{Young:2008aa} that integrated high content
 screening and ligand-target prediction models to construct network
@@
 Networks constructed from drugs and their protein targets were first
-described by Yildirim et al \cite{Yildirim:2007aa}, who considered
+described by Yildirim et al. \cite{Yildirim:2007aa}, who considered
 approved and investigational drugs. Networks built using the former
@@
 still an open question as to whether such predicted networks can serve
-as the foundation for downstream analyses, though Pahikkala et al
+as the foundation for downstream analyses, though Pahikkala et al. 
 \cite{Pahikkala:2014gt} describe a number of guidelines to enhance the
@@
 context of known protein-protein and protein-compound
-interactions. Iwata et al \cite{Iwata:2013lk} described the use of a
+interactions. Iwata et al. \cite{Iwata:2013lk} described the use of a
 drug-target network to identify protein domains that are associated
-with drug side effects. Wang et al \cite{Wang:2013gn} characterized
+with drug side effects. Wang et al. \cite{Wang:2013gn} characterized
 the topological properties of drug-target networks and noted that
@@
 predictive models. For example, a model to predict ATC codes from a
-drug-target network was described by Wang et al
-\cite{Wang:2013ye}. Tabei et al \cite{Tabei:2012tn} used a drug-target
+drug-target network was described by Wang et al. 
+\cite{Wang:2013ye}. Tabei et al. \cite{Tabei:2012tn} used a drug-target
 network as input to a classifier that was used to extract chemogenomic
@@
 A number of workers have used drug-target network information as the
-basis of drug repositioning methodologies. Wang et al
+basis of drug repositioning methodologies. Wang et al.
 \cite{Wang:2014lp} who combined drug-target and disease-target
-networks to generate repositioning hypotheses. Tan et al
+networks to generate repositioning hypotheses. Tan et al. 
 \cite{Tan:2014qb} employed drug-target networks as a component of a
@@
 semantic similarity to identify new indications for a number of
-preexisting drugs.  Peng et al \cite{Peng:2014vy} considered a variant
-of the drug-target network concept by developing an interaction
-network of FDA approved drugs and protein pockets. Drugs were docked
-into each pocket and the docking score was used to define edges
-between the drug and protein whose pocket it was docked to.
+preexisting drugs.  Peng et al. \cite{Peng:2014vy} considered a
+variant of the drug-target network concept by developing an
+interaction network of FDA approved drugs and protein pockets. Drugs
+were docked into each pocket and the docking score was used to define
+edges between the drug and protein whose pocket it was docked to.
 
@@
 
-Schadt et al \cite{Schadt:2009zl} have stressed the importance of
+Schadt et al. \cite{Schadt:2009zl} have stressed the importance of
 viewing diseases as manifestations of perturbed biological networks
@@
 different classes of entities. An example is the network constructed
-from long non-coding RNAs and diseases by Zhou et al
+from long non-coding RNAs and diseases by Zhou et al. 
 \cite{Zhou:2015if}. Based on this network, a random walk approach was
@@
 opportunities for repurposing drugs based on novel disease-target or
-disease-pathway associations. For example, Pacini et al
+disease-pathway associations. For example, Pacini et al. 
 \cite{Pacini:2013kl} describe DvD, a pipeline comprised of R and
@@
 affect multiple targets and pathways. This approach was exemplified by
-Lehar et al \cite{Lehar:2009gu} who proposed that drug combination
+Lehar et al. \cite{Lehar:2009gu} who proposed that drug combination
 response surfaces were functionally related to the underlying
@@
 Network similarity graphs (NSG's) are an alternative network
-visualization of SAR data described by Wawer et al \cite{Wawer:2008aa}
+visualization of SAR data described by Wawer et al. \cite{Wawer:2008aa}
 that employs a different characterization of activity cliffs. This
@@
 Networks have been applied to a number of other SAR related
-problems. For example, Webb et al \cite{Webb:2014tp} proposed a
+problems. For example, Webb et al. \cite{Webb:2014tp} proposed a
 network approach to the interpretation of arbitrary machine learning
-models. Hanser et al \cite{Hanser:2014gl} constructed a network model
+models. Hanser et al. \cite{Hanser:2014gl} constructed a network model
 to represent SAR knowledge data, resulting in a ``hypothesis
@@
 \href{http://swami.wustl.edu/flow}{http://swami.wustl.edu/flow}.
-Swamidass et al \cite{Swamidass:2014vn} describe an approach to
+Swamidass et al. \cite{Swamidass:2014vn} describe an approach to
 constructing ``assay networks'' based on Pubchem data, where assays
@@
 
-Varin et al \cite{Varin:2011ve} describe ``scaffold networks'' which
+Varin et al. \cite{Varin:2011ve} describe ``scaffold networks'' which
 are similar in nature to scaffold trees but instead of a tree
@@
   scaffolds are generated only by the Scaffold Network
-  method. Modified from Figure 1 in Varin et al \cite{Varin:2011ve}.}
+  method. Modified from Figure 1 in Varin et al. \cite{Varin:2011ve}.}
   \label{fig:scaffnet}
@@
 drug-disease (or even drug-disease-target) networks are examples of
-such approaches. Hartsperger et al \cite{Hartsperger:2010yg} describe
+such approaches. Hartsperger et al. \cite{Hartsperger:2010yg} describe
 an approach to fuzzy clustering that is capable of identifying
@@
 by random walk approaches on bipartite networks such as that described
-by Chen et al \cite{Chen:2012qy} whose proposed a random walk approach
+by Chen et al. \cite{Chen:2012qy} whose proposed a random walk approach
 to predict drug-target interactions.

### 2015-03-22 paper.tex
updated figure caption
minor fix to spacing

@@
 by integrating multiple data types. An example of this was described
-by Young et al. \cite{Young:2008aa} that integrated high content
+by Young et al.~\cite{Young:2008aa} that integrated high content
 screening and ligand-target prediction models to construct network
@@
 Networks constructed from drugs and their protein targets were first
-described by Yildirim et al. \cite{Yildirim:2007aa}, who considered
+described by Yildirim et al.~\cite{Yildirim:2007aa}, who considered
 approved and investigational drugs. Networks built using the former
@@
 still an open question as to whether such predicted networks can serve
-as the foundation for downstream analyses, though Pahikkala et al. 
+as the foundation for downstream analyses, though Pahikkala et al.~
 \cite{Pahikkala:2014gt} describe a number of guidelines to enhance the
@@
 context of known protein-protein and protein-compound
-interactions. Iwata et al. \cite{Iwata:2013lk} described the use of a
+interactions. Iwata et al.~\cite{Iwata:2013lk} described the use of a
 drug-target network to identify protein domains that are associated
-with drug side effects. Wang et al. \cite{Wang:2013gn} characterized
+with drug side effects. Wang et al.~\cite{Wang:2013gn} characterized
 the topological properties of drug-target networks and noted that
@@
 predictive models. For example, a model to predict ATC codes from a
-drug-target network was described by Wang et al. 
-\cite{Wang:2013ye}. Tabei et al. \cite{Tabei:2012tn} used a drug-target
+drug-target network was described by Wang et al.~
+\cite{Wang:2013ye}. Tabei et al.~\cite{Tabei:2012tn} used a drug-target
 network as input to a classifier that was used to extract chemogenomic
@@
 \cite{Wang:2014lp} who combined drug-target and disease-target
-networks to generate repositioning hypotheses. Tan et al. 
+networks to generate repositioning hypotheses. Tan et al.~
 \cite{Tan:2014qb} employed drug-target networks as a component of a
@@
 semantic similarity to identify new indications for a number of
-preexisting drugs.  Peng et al. \cite{Peng:2014vy} considered a
+preexisting drugs.  Peng et al.~\cite{Peng:2014vy} considered a
 variant of the drug-target network concept by developing an
@@
 
-Schadt et al. \cite{Schadt:2009zl} have stressed the importance of
+Schadt et al.~\cite{Schadt:2009zl} have stressed the importance of
 viewing diseases as manifestations of perturbed biological networks
@@
 different classes of entities. An example is the network constructed
-from long non-coding RNAs and diseases by Zhou et al. 
+from long non-coding RNAs and diseases by Zhou et al.~
 \cite{Zhou:2015if}. Based on this network, a random walk approach was
@@
 opportunities for repurposing drugs based on novel disease-target or
-disease-pathway associations. For example, Pacini et al. 
+disease-pathway associations. For example, Pacini et al.~
 \cite{Pacini:2013kl} describe DvD, a pipeline comprised of R and
@@
 affect multiple targets and pathways. This approach was exemplified by
-Lehar et al. \cite{Lehar:2009gu} who proposed that drug combination
+Lehar et al.~\cite{Lehar:2009gu} who proposed that drug combination
 response surfaces were functionally related to the underlying
@@
   \includegraphics[width=\linewidth]{img/salinet} 
-  \caption{A SALI network constructed from XXX. See Ref.~\cite{Guha:2008aa} for further details.}
+  \caption{A SALI network constructed from set of 62 dihydroquinoline
+    derivatives that were designed to inhibit the glucocorticoid
+    receptor. See Ref.~\cite{Guha:2008aa} for further details.}
   \label{fig:salinet}
@@
 Network similarity graphs (NSG's) are an alternative network
-visualization of SAR data described by Wawer et al. \cite{Wawer:2008aa}
+visualization of SAR data described by Wawer et al.~\cite{Wawer:2008aa}
 that employs a different characterization of activity cliffs. This
@@
 Networks have been applied to a number of other SAR related
-problems. For example, Webb et al. \cite{Webb:2014tp} proposed a
+problems. For example, Webb et al.~\cite{Webb:2014tp} proposed a
 network approach to the interpretation of arbitrary machine learning
-models. Hanser et al. \cite{Hanser:2014gl} constructed a network model
+models. Hanser et al.~\cite{Hanser:2014gl} constructed a network model
 to represent SAR knowledge data, resulting in a ``hypothesis
@@
 \href{http://swami.wustl.edu/flow}{http://swami.wustl.edu/flow}.
-Swamidass et al. \cite{Swamidass:2014vn} describe an approach to
+Swamidass et al.~\cite{Swamidass:2014vn} describe an approach to
 constructing ``assay networks'' based on Pubchem data, where assays
@@
 
-Varin et al. \cite{Varin:2011ve} describe ``scaffold networks'' which
+Varin et al.~\cite{Varin:2011ve} describe ``scaffold networks'' which
 are similar in nature to scaffold trees but instead of a tree
@@
   scaffolds are generated only by the Scaffold Network
-  method. Modified from Figure 1 in Varin et al. \cite{Varin:2011ve}.}
+  method. Modified from Figure 1 in Varin et al.~\cite{Varin:2011ve}.}
   \label{fig:scaffnet}
@@
 drug-disease (or even drug-disease-target) networks are examples of
-such approaches. Hartsperger et al. \cite{Hartsperger:2010yg} describe
+such approaches. Hartsperger et al.~\cite{Hartsperger:2010yg} describe
 an approach to fuzzy clustering that is capable of identifying
@@
 by random walk approaches on bipartite networks such as that described
-by Chen et al. \cite{Chen:2012qy} whose proposed a random walk approach
+by Chen et al.~\cite{Chen:2012qy} whose proposed a random walk approach
 to predict drug-target interactions.
